---
title: "Convergence rates for the interval graph"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography: 
  - ../references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Go back to the [Convergence Rates page](../convergence_rates.html).

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  #out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
# Install R-INLA package
# install.packages("INLA",repos = c(getOption("repos"),INLA ="https://inla.r-inla-download.org/R/testing"), dep = TRUE)
# Update R-INLA package
# inla.upgrade(testing = TRUE)
# Install inlabru package
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# Install rSPDE package
# remotes::install_github("davidbolin/rspde", ref = "devel")
# Install MetricGraph package
# remotes::install_github("davidbolin/metricgraph", ref = "devel")

library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)
library(Matrix)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)
library(ggplot2)
library(reshape2)
library(sf)

library(here)
library(rmarkdown)
library(knitr)
library(grateful) # Cite all loaded packages

library(latex2exp)
library(plotrix)

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```


# Interval graph

## Define utility functions

```{r}
# Matern covariance function
matern.covariance <- function(h, kappa, nu, sigma) {
  if (nu == 1 / 2) {
    C <- sigma^2 * exp(-kappa * abs(h))
  } else {
    C <- (sigma^2 / (2^(nu - 1) * gamma(nu))) *
      ((kappa * abs(h))^nu) * besselK(kappa * abs(h), nu)
  }
  C[h == 0] <- sigma^2
  return(as.matrix(C))
}

# Folded.matern.covariance.1d
folded.matern.covariance.1d.local <- function(x, kappa, nu, sigma,
                                              L = 1, N = 10,
                                              boundary = c("neumann",
                                                           "dirichlet", "periodic")) {
  boundary <- tolower(boundary[1])
  if (!(boundary %in% c("neumann", "dirichlet", "periodic"))) {
    stop("The possible boundary conditions are 'neumann',
    'dirichlet' or 'periodic'!")
  }
  addi <- t(outer(x, x, "+"))
  diff <- t(outer(x, x, "-"))
  s1 <- sapply(-N:N, function(j) { 
    diff + 2 * j * L
  })
  s2 <- sapply(-N:N, function(j) {
    addi + 2 * j * L
  })
  if (boundary == "neumann") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) +
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else if (boundary == "dirichlet") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) -
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else {
    C <- rowSums(matern.covariance(h = s1,
                                   kappa = kappa, nu = nu, sigma = sigma))
  }
  return(matrix(C, nrow = length(x)))
}

# Function to get the true covariance matrix
gets_true_cov_mat = function(graph, kappa, nu, sigma, N, boundary){
  h <- graph$mesh$V[,1]
  true_cov_mat <- folded.matern.covariance.1d.local(x = h, kappa = kappa, nu = nu, sigma = sigma, N = N, boundary = boundary)
  return(true_cov_mat)
}
```

## Define the graph

```{r, fig.dim = c(16,8)}
edge <- rbind(c(0,0),c(1,0))
edges <- list(edge)
graph <- metric_graph$new(edges = edges)
graph$plot() + 
  ggtitle("Interval graph") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino")) +
  coord_fixed(ratio = 2)
```

## Define the parameters

```{r}
# parameters
h.ok <- 2^-9
type <- "covariance"
type_rational_approximation = "brasil"
rho <- 0.5
#m = 4
sigma <- 1
N.folded <- 10
boundary <- "neumann" # do not change this

# Mesh sizes
h_aux <- c(6:3)
h_vector <- 2^-h_aux
h_label <- paste0("2^-", h_aux, "")
h_label_latex <- sprintf("$2^{-%d}$", h_aux)

# Beta values
beta_aux <- c(4, 4.5, 5, 5.5, 6, 6.5, 7)
beta_vector <- beta_aux/8
beta_label <- paste0(beta_aux, "/8")
theoretical_rate <- pmin(4*beta_vector-1/2,2)
```

## Build the graph with overkill mesh

```{r, fig.dim = c(16,8)}
graph.ok <- graph$clone()
# Build graph with overkill mesh
graph.ok$build_mesh(h = h.ok)
graph.ok$compute_fem()

graph.ok$plot(mesh = TRUE) + 
  ggtitle("Interval graph with overkill mesh") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino")) +
  coord_fixed(ratio = 2)
```


```{r}
# Get the overkill mesh locations
loc.ok <- graph.ok$mesh$VtE # or graph.ok$get_mesh_locations()
```


Each `A[[i]]` below is the projection matrix corresponding to the graph with mesh size `h_vector[i]` onto the overkill mesh.

```{r}
# Initialize the list of graphs and the list of projection matrices
graphs <- list()
A <- list()
for(i in 1:length(h_vector)){
  graphs[[i]] <- graph$clone()
  graphs[[i]]$build_mesh(h = h_vector[i])
  A[[i]] <- graphs[[i]]$fem_basis(loc.ok)
}
# Print the dimensions of the projection matrices
print(lapply(A, dim))
```

## Compute the covariance error

```{r, fig.dim = c(16, 20)}
cov.error.ok.mesh <- matrix(NA, nrow = length(h_vector), ncol = length(beta_vector))
cov.error.folded <- matrix(NA, nrow = length(h_vector), ncol = length(beta_vector))

par(family = "Palatino")
layout(matrix(1:(length(beta_vector)*length(h_vector)), 
              nrow = length(beta_vector), 
              byrow = TRUE))    

m_values <- c()
for (j in 1:length(beta_vector)) {
  beta <- beta_vector[j]
  alpha <- 2*beta
  fract2beta <- alpha - floor(alpha)
  nu <- alpha - 0.5
  kappa <- sqrt(8*nu)/rho
  tau <- sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
  
  Sigma.folded <- gets_true_cov_mat(graph = graph.ok, # using folded
                               kappa = kappa,
                               nu = nu,
                               sigma = sigma,
                               N = N.folded,
                               boundary = boundary)
  
  for (i in 1:length(h_vector)) {
    h <- h_vector[i]
    m <- min(20, ceiling((min(4*beta - 1/2,2) + 1/2)^2*log(h)^2/(4*pi^2*fract2beta)))
    m_values <- c(m_values, m)
    
    Sigma.ok.mesh <- matern.operators(alpha = alpha, # using overkill mesh
                                kappa = kappa,
                                tau = tau,
                                m = m,
                                graph = graph.ok,
                                type = type,
                                type_rational_approximation = type_rational_approximation)$covariance_mesh()
    
    
    Sigma <- matern.operators(alpha = alpha,  # this is the approximation
                             kappa = kappa, 
                             tau = tau,
                             m = m, 
                             graph = graphs[[i]],
                             type = type,
                             type_rational_approximation = type_rational_approximation)$covariance_mesh()
    
    Sigma.approx <- A[[i]]%*%Sigma%*%t(A[[i]])
    
    cov.error.ok.mesh[i,j] <- sqrt(as.double(t(graph.ok$mesh$weights)%*%(Sigma.ok.mesh - Sigma.approx)^2%*%graph.ok$mesh$weights))
    cov.error.folded[i,j] <- sqrt(as.double(t(graph.ok$mesh$weights)%*%(Sigma.folded - Sigma.approx)^2%*%graph.ok$mesh$weights))
 
    aux <- dim(Sigma.approx)[1]
    reordering <- c(1,3:aux,2)
    auxhalf <- reordering[length(reordering) %/% 2 + 1]

    plot(Sigma.folded[auxhalf, reordering], col = "black", type = "l", main = bquote(beta == .(beta_label[j]) ~ "," ~ h == .(h_label[i])), xlab = "", ylab = "", lwd = 2, lty = 1)
    lines(Sigma.ok.mesh[auxhalf, reordering], col = "red", lty = 2, lwd = 2)
    lines(Sigma.approx[auxhalf, reordering], col = "blue", lty = 3, lwd = 2)
    legend("topleft", lwd = 2, col = c("black", "red", "blue"), legend = c("true", "ok.mesh", "approx"), lty = c(1,2,3))
  }
}
print(m_values)
```

```{r}
# theoretical
theoretical_rate
# folded
for (u in 1:length(beta_vector)) {print(coef(lm(log(cov.error.folded[,u]) ~ log(h_vector))))}
# ok.mesh
for (u in 1:length(beta_vector)) {print(coef(lm(log(cov.error.ok.mesh[,u]) ~ log(h_vector))))}
```


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
loglog_line_equation <- function(x1, y1, slope) {
  b <- log10(y1 / (x1 ^ slope))
  
  function(x) {
    (x ^ slope) * (10 ^ b)
  }
}

guiding_lines <- matrix(NA, nrow = length(h_vector), ncol = length(beta_vector))
for (j in 1:length(beta_vector)) {
  guiding_lines_aux <- matrix(NA, nrow = length(h_vector), ncol = length(h_vector))
  for(k in 1:length(h_vector)){
    point_x1 <- h_vector[k]
    point_y1 <- cov.error.ok.mesh[k, j]
    slope <- theoretical_rate[j]
    line <- loglog_line_equation(x1 = point_x1, y1 = point_y1, slope = slope)
    guiding_lines_aux[,k] <- line(h_vector)
  }
  guiding_lines[,j] <- apply(guiding_lines_aux, 1, mean)
}

# Generate default ggplot2 colors
default_colors <- scales::hue_pal()(ncol(guiding_lines))

# Create the plot_lines list with different colors for each line
plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
  geom_line(data = data.frame(x = h_vector, y = guiding_lines[, i]),
            aes(x = x, y = y), color = default_colors[i], linetype = "dashed", show.legend = FALSE)
})

df <- as.data.frame(cbind(h_vector, cov.error.ok.mesh))
colnames(df) <- c("h_vector", beta_label)
df_melted <- melt(df, id.vars = "h_vector", variable.name = "column", value.name = "value")

p <- ggplot() +
  geom_line(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  geom_point(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  plot_lines +
  labs(title = "Covariance error using overkill mesh",
    x = TeX("$\\hat{h}$"),
       y = "Covariance Error",
       color = expression(beta)) +
  scale_x_log10(breaks = h_vector, labels = TeX(h_label_latex)) +
  scale_y_log10() +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))
```


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
guiding_lines <- matrix(NA, nrow = length(h_vector), ncol = length(beta_vector))
for (j in 1:length(beta_vector)) {
  guiding_lines_aux <- matrix(NA, nrow = length(h_vector), ncol = length(h_vector))
  for(k in 1:length(h_vector)){
    point_x1 <- h_vector[k]
    point_y1 <- cov.error.folded[k, j]
    slope <- theoretical_rate[j]
    line <- loglog_line_equation(x1 = point_x1, y1 = point_y1, slope = slope)
    guiding_lines_aux[,k] <- line(h_vector)
  }
  guiding_lines[,j] <- apply(guiding_lines_aux, 1, mean)
}

# Generate default ggplot2 colors
default_colors <- scales::hue_pal()(ncol(guiding_lines))

# Create the plot_lines list with different colors for each line
plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
  geom_line(data = data.frame(x = h_vector, y = guiding_lines[, i]),
            aes(x = x, y = y), color = default_colors[i], linetype = "dashed", show.legend = FALSE)
})

df <- as.data.frame(cbind(h_vector, cov.error.folded))
colnames(df) <- c("h_vector", beta_label)
df_melted <- melt(df, id.vars = "h_vector", variable.name = "column", value.name = "value")

q <- ggplot() +
  geom_line(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  geom_point(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  plot_lines +
  labs(title = "Covariance error using folded",
    x = TeX("$\\hat{h}$"),
       y = "Covariance Error",
       color = expression(beta)) +
  scale_x_log10(breaks = h_vector, labels = TeX(h_label_latex)) +
  scale_y_log10() +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))
```

```{r, fig.dim = c(12, 5)}
p + q
```



# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
