---
title: "Functions on metric graphs"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---


```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Go back to the [About page](about.html).

Go back to the [Basis Functions](basis_functions.html) page.

Let us set some global options for all code chunks in this document.


```{r}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

# Import libraries

```{r}
library(MetricGraph)
library(dplyr)
library(plotly)

library(here)
library(rmarkdown)
library(grateful) # to cite
```


# Building the graph object

```{r}
V <- rbind(c(0, 0), c(1, 0), c(1, 1), c(0, 1), c(-1, 1), c(-1, 0), c(0, -1))
E <- rbind(c(1, 2), c(2, 3), c(3, 4), c(4, 5),
           c(5, 6), c(6, 1), c(4, 1),c(1, 7))
graph <- metric_graph$new(V = V, E = E)
```

# Geodesic distance on a metric graph {#geodesic}

Figure 1 below shows an illustration of a compact and connected graph $\Gamma = (\mathcal{V},\mathcal{E})$, where $\mathcal{E}= \{e_1,\dots,e_8\}$ (black lines) and $\mathcal{V} =  \{v_1,\dots,v_7\}$ (blue points).


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, fig.height = 8, out.width = "100%", class.source = "fold-hide", fig.cap = captioner("Illustration of a compact and connected graph $\\Gamma = \\pare{\\Vcal,\\Ecal}$, where $\\Ecal= \\llav{e_1,\\dots,e_8}$ (black) and $\\Vcal =  \\llav{v_1,\\dots,v_7}$ (blue). Notice that $\\ell_{e_i} = 1$ for $i = 1,\\dots,8$. Note also that, for example, $\\Ecal_{v_2} = \\llav{e_1,e_2,e_5,e_8}$ and therefore $d_{v_2} = 4.$ Given two points $s_1 = (e_3, 0.5)$ and $s_2 = (e_1, 0.5)$, we have that $d(s_1,s_2) = 2$ (red). Following the convention mentioned in the text, we may write, for example, $e_8 = (v_2,v_7) \\cong [0,\\ell_{e_8}]$.")}
df = data.frame(x = c(0.5,0), y = c(-1,-1), z = c(0,0))
df2 = data.frame(x = c(0,0), y = c(-1,0), z = c(0,0))
df3 = data.frame(x = c(0,-0.5), y = c(0,0), z = c(0,0))

graph$build_mesh(h = 1/40)
x <- graph$mesh$V[, 1]
y <- graph$mesh$V[, 2]
f <- x^2 - y^2
graph$plot(plotly = TRUE, vertex_size = 5, vertex_color = "blue",
           edge_color = "black", edge_width = 2) %>%  
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\text{Geodesic distance on a metric graph }\\Gamma = (\\mathcal{V},\\mathcal{E})"), 
    showlegend = FALSE,
                    font = list(family = "Palatino"),
                    scene = list(
                      aspectratio = list(x = 1.6, y = 1.6, z = 1),
                      annotations = list(
                        list(
                          x = -1, y = 0, z = 0,
                          text = TeX("v_1"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 0, z = 0,
                          text = TeX("v_2"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = -1, z = 0,
                          text = TeX("v_3"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = -1, z = 0,
                          text = TeX("v_4"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 0, z = 0,
                          text = TeX("v_5"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 1, z = 0,
                          text = TeX("v_6"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 1, z = 0,
                          text = TeX("v_7"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = -1, z = 0,
                          text = TeX("s_1"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = -0.5, y = 0, z = 0,
                          text = TeX("s_2"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 20),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = -0.5, y = 0, z = 0,
                          text = TeX("e_1"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = -0.5, z = 0,
                          text = TeX("e_2"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = -1, z = 0,
                          text = TeX("e_3"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = -0.5, z = 0,
                          text = TeX("e_4"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = 0, z = 0,
                          text = TeX("e_5"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 0.5, z = 0,
                          text = TeX("e_6"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = 1, z = 0,
                          text = TeX("e_7"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 0.5, z = 0,
                          text = TeX("e_8"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 20),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1)
                        ))) %>% 
  add_trace(x = 0.5, y = -1, z = 0, mode = "markers", type = "scatter3d", 
            marker = list(size = 5, color = "red", symbol = 104)) %>% 
  add_trace(x = -0.5, y = 0, z = 0, mode = "markers", type = "scatter3d", 
            marker = list(size = 5, color = "red", symbol = 104)) %>% 
  add_trace(data = df, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104)) %>% 
  add_trace(data = df2, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104)) %>% 
  add_trace(data = df3, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104))
```


# Function on a metric graph {#functions-def-on-graph}

Figure 2 below provides an illustration of a function defined on a graph.

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, fig.height = 8, out.width = "100%", class.source = "fold-hide", fig.cap = captioner("Illustration of a function $f$ (in blue) defined on the graph $\\Gamma= (\\Vcal,\\Ecal)$, where $\\Ecal= \\llav{e_1,\\dots,e_8}$ and $\\Vcal =  \\llav{v_1,\\dots,v_7}$. Function $f:\\Gamma\\longrightarrow\\ar$ is the  restriction of the surface $\\gamma:\\ar^2\\longrightarrow\\ar$ to $\\Gamma$, where $\\gamma(x,y) = x^2-y^2$. That is, $f= \\gamma\\big|_\\Gamma$.")}
graph$build_mesh(h = 1/40)
x <- graph$mesh$V[, 1]
y <- graph$mesh$V[, 2]
f <- x^2 - y^2

df = data.frame(x = c(0.5,0), y = c(-1,-1), z = c(0,0))
df2 = data.frame(x = c(0,0), y = c(-1,0), z = c(0,0))
df3 = data.frame(x = c(0,-0.5), y = c(0,0), z = c(0,0))


graph$plot_function(f,plotly = TRUE, vertex_size = 5, vertex_color = "blue", line_color = "blue",
           edge_color = "black", edge_width = 2) %>%  
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\text{Function on a metric graph }\\Gamma = (\\mathcal{V},\\mathcal{E})"), 
    showlegend = FALSE,
                    font = list(family = "Palatino"),
                    scene = list(
                      aspectratio = list(x = 1.6, y = 1.6, z = 1),
                      annotations = list(
                        list(
                          x = -1, y = 0, z = 0,
                          text = TeX("v_1"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 0, z = 0,
                          text = TeX("v_2"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = -1, z = 0,
                          text = TeX("v_3"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = -1, z = 0,
                          text = TeX("v_4"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 0, z = 0,
                          text = TeX("v_5"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 1, z = 0,
                          text = TeX("v_6"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 1, z = 0,
                          text = TeX("v_7"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = -1, z = 0,
                          text = TeX("s_1"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = -0.5, y = 0, z = 0,
                          text = TeX("s_2"),
                          textangle = 0, ax = 0, ay = -25,
                          font = list(color = "black", size = 16),
                          arrowcolor = "black", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = -0.5, y = 0, z = 0,
                          text = TeX("e_1"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = -0.5, z = 0,
                          text = TeX("e_2"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = -1, z = 0,
                          text = TeX("e_3"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = -0.5, z = 0,
                          text = TeX("e_4"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = 0, z = 0,
                          text = TeX("e_5"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 1, y = 0.5, z = 0,
                          text = TeX("e_6"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0.5, y = 1, z = 0,
                          text = TeX("e_7"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = 0.5, z = 0,
                          text = TeX("e_8"),
                          textangle = 0, ax = 0, ay = 10,
                          font = list(color = "black", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1),
                        list(
                          x = 0, y = -1, z = 1,
                          text = TeX("f=\\gamma\\big|_\\Gamma"),
                          textangle = 0, ax = 60, ay = 0,
                          font = list(color = "blue", size = 16),
                          arrowcolor = "white", arrowsize = 1, arrowwidth = 1,arrowhead = 1)
                        ))) %>% 
  add_trace(x = 0.5, y = -1, z = 0, mode = "markers", type = "scatter3d", 
            marker = list(size = 5, color = "red", symbol = 104)) %>% 
  add_trace(x = -0.5, y = 0, z = 0, mode = "markers", type = "scatter3d", 
            marker = list(size = 5, color = "red", symbol = 104)) %>% 
  add_trace(data = df, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104)) %>% 
  add_trace(data = df2, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104)) %>% 
  add_trace(data = df3, x = ~x, y = ~y, z = ~z, mode = "lines", type = "scatter3d",
            line = list(width = 2, color = "red", symbol = 104))
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
