---
title: "Computing the marginal standard deviation explicitly and computationally efficiently"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Below we set some global options for all code chunks in this document.


```{r}
rm(list = ls())
# Set seed for reproducibility
set.seed(593) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```


Below we load the necessary libraries and define the auxiliary functions.

```{r}
library(rSPDE)
library(MetricGraph)
library(sp)
library(osmdata)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)

library(here)
library(rmarkdown)
# Cite all loaded packages
library(grateful)
```



```{r}
rtilde_matern_interval <- function(kappa, tau, le, t1, t2) (1 / (2 * kappa * tau^2 * sinh(kappa * le))) * (cosh(kappa * (le - abs(t1 - t2))) + cosh(kappa * (t1 + t2 - le)))
sparse_A <- function(n) bandSparse(n-1, n, k = c(0, 1), diagonals = list(rep(1, n), rep(-1, n - 1)))
```

Let's build a small graph to illustrate how to compute the marginal standard deviation.


```
{r}
edge1 <- rbind(c(0,0), c(1,0))[2:1, ]
theta <- seq(from = -pi, to = pi,length.out = 100)
edge2 <- cbind(1+1/pi+cos(theta)/pi, sin(theta)/pi)
edge3 <- rbind(c(-1,-1), c(1,0))[2:1, ]
edges <- list(edge1, edge2, edge3)
graph <- metric_graph$new(edges = edges)
```
```
{r}
edge <- rbind(c(0,0),c(1,0))
edges <- list(edge)
graph <- metric_graph$new(edges = edges)
```

```
{r}
edges_roots_of_unity <- function(n, length = 10) {
  x_coords <- cos(2 * pi * (0:(n-1)) / n)
  y_coords <- sin(2 * pi * (0:(n-1)) / n)
  return(lapply(1:n, function(i) {rbind(c(0, 0), c(length*x_coords[i], length*y_coords[i]))}))
}
n <- 10
my_length <- 2
graph <- metric_graph$new(edges = edges_roots_of_unity(n, length = my_length))
```

```{r}
a <- 3
b <- 12
edge2_x <- runif(1, 8, 12)

edge1 <- rbind(c(0,0),c(-runif(1, a, b),0))
edge2 <- rbind(c(0,0),c(edge2_x,0))
edge3 <- rbind(c(edge2_x,0),c(edge2_x + runif(1, a, b),0))
edge4 <- rbind(c(edge2_x,0),c(edge2_x + runif(1, a, b),-runif(1, a, b)))
edge5 <- rbind(c(edge2_x,0),c(edge2_x + runif(1, a, b),runif(1, a, b)))
edge6 <- rbind(c(0,0),c(-runif(1, a, b),-runif(1, a, b)))
edge7 <- rbind(c(0,0),c(-runif(1, a, b), runif(1, a, b)))

edges <- list(edge1, edge2, edge3, edge4, edge5, edge6, edge7)
graph <- metric_graph$new(edges = edges)
```

```
{r}
edges <- logo_lines() 
graph <- metric_graph$new(edges = edges, perform_merges = TRUE) 
```

```
{r}
call <- opq(bbox = c(39.0884, 22.33, 39.115, 22.3056))
call <- add_osm_feature(call, key = "highway",value=c("motorway",
                                                      "primary","secondary",
                                                      "tertiary",
                                                      "residential"))
data <- osmdata_sf(call)
graph <- graph_components$new(data)
graph <- graph$get_largest()
graph$prune_vertices()
```


```{r}
graph_copy <- graph$clone()
graph_pcr_1 <- graph$clone()
graph_pcr_2 <- graph$clone()
graph$build_mesh(h = 0.1)
```


```{r}
sigma <- 1
range <- 1
nu <- 1/2

kappa <- sqrt(8*nu)/range
tau <-sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))
alpha <- nu + 1/2 #alpha = 1

rtilde_int <- function(t1, t2, le) rtilde_matern_interval(kappa = kappa, tau = tau, le, t1, t2)
```



```{r}
get_marginal_std <- function(kappa, tau, graph){
  
  V <- graph$V
  E <- graph$E
  degrees <- graph$get_degrees() 
  edge_lengths <- as.vector(graph$get_edge_lengths())
  VtE <- graph$mesh$VtE
  
  
  if(nrow(V) == 2 & nrow(E) == 1){
    aux_function <- function(VtErow) {
      edge <- VtErow[1]
      s <- VtErow[2]
      le <- edge_lengths[edge]
      return(rtilde_int(le*s, le*s, le))
  }
  } else {
    # Build matrix U
    U <- matrix(nrow = 0, ncol = 2)
    for (vertex in seq_len(nrow(V))) {
      if (degrees[vertex] > 1) {
        edges_leaving <- cbind(which(E[,1] %in% vertex), 0) 
        edges_entering <- cbind(which(E[,2] %in% vertex), 1) 
        if (sum(dim(edges_leaving)) == 2 ) {edges_leaving <- NULL}
        if (sum(dim(edges_entering)) == 2 ) {edges_entering <- NULL}
        constraints <- rbind(edges_leaving, edges_entering)
        ordered_constraints <- constraints[order(constraints[, 1], constraints[, 2]),]
        U <- rbind(U, ordered_constraints)
      }
    }
    # Build matrix A
    A <- bdiag(lapply(degrees[degrees >1], sparse_A))
    # Build matrix Sigma
    edge_lengths_in_U <- edge_lengths[U[, 1]] 
    s_diag <- edge_lengths_in_U*U[, 2] 
    diag <- as.vector(mapply(rtilde_int, t1 = s_diag, t2 = s_diag, le = edge_lengths_in_U))
    Sigma <- Diagonal(x = diag)
    for (edge in seq_len(nrow(E))) {
      pos <- which(U[, 1] == edge)
      if (length(pos) > 1) {
        Sigma[pos[1], pos[2]] <- Sigma[pos[2], pos[1]] <- rtilde_int(t1 = 0, t2 = edge_lengths[edge], le = edge_lengths[edge])
      }
    }
  
    ASigmaAt <- A %*% Sigma %*% t(A)
    
    aux_function <- function(VtErow) {
      edge <- VtErow[1]
      s <- VtErow[2]
      le <- edge_lengths[edge]
      pos <- which(U[, 1] == edge)
      cc <- mapply(rtilde_int, t1 = le*s, t2 = U[pos,2]*le, le = le)
      v <- if(length(pos) == 1) as.vector(A[, pos] * cc) else as.vector(A[, pos] %*% cc)
      return(rtilde_int(le*s, le*s, le) - t(v) %*% solve(ASigmaAt, v))
    }
  }
  return(sqrt(apply(X = VtE, MARGIN = 1, FUN = aux_function)))
}
```

```{r}
compute_pcr_2v <- function(cov_matrix, graph, cor_threshold = 0.1) {
  
  # Compute the correlation matrix
  cor_matrix <-cov2cor(cov_matrix)
  # Compute the geodesic distance matrix
  graph$compute_geodist_mesh() 
  # Extract the geodesic distance matrix
  dist_matrix <- graph$mesh$geo_dist 
  # Initialize the vector to store the practical correlation range
  pcr <- numeric(dim(cor_matrix)[1]) 
  
  process_row <- function(row_index) {
    # For each row_index, create an auxiliary matrix aux_mat with the correlation and geodesic distance
    aux_mat <- cbind(cor_matrix[row_index, ], dist_matrix[row_index, ]) 
    # Order the auxiliary matrix by the geodesic distance
    ordered_aux_mat <- aux_mat[order(aux_mat[, 2]), ] 
    # Filter the auxiliary ordered matrix by the correlation threshold
    filtered_aux_mat <- ordered_aux_mat[ordered_aux_mat[,1] < cor_threshold, ] 
    if (nrow(filtered_aux_mat) > 0) {
      # Return the minimum geodesic distance such that the correlation is below the threshold
      return(filtered_aux_mat[1, 2]) 
    } else {
      # If the condition is not satisfied, record the minimum correlation value and return an error message
      min <- min(aux_mat[, 1])
      stop(paste0("The condition cor_matrix[row_index, ] < cor_threshold is not satisfied for row_index = ", 
                  row_index, ". Increase cor_threshold. The minimum value of cor_matrix[row_index, ] is ",
                  min))
    }
  }
  
  pcr <- sapply(1:dim(cor_matrix)[1], process_row)
  
  return(pcr)
}
```


```{r}
std_cheap <- get_marginal_std(kappa = kappa, tau = tau, graph = graph)
```

```{r}
mesh_loc <- graph$get_mesh_locations() %>% 
  as.data.frame() %>% 
  mutate(y = 1) %>% 
  rename(edge_number = V1, distance_on_edge = V2)

graph_copy$add_observations(mesh_loc,
                       edge_number = "edge_number",
                       distance_on_edge = "distance_on_edge",
                       data_coords = "PtE",
                       normalized = TRUE, 
                       clear_obs = TRUE)
graph_copy$observation_to_vertex()

Q <- spde_precision(kappa = kappa,
               tau = tau,
               alpha = alpha,
               graph = graph_copy,
               BC = 0) # BC = 1

Sigma <- solve(Q)
std_expensive <- sqrt(diag(Sigma))

nonstat_cov <-  Matrix::Diagonal(x = 1/std_cheap)%*% Sigma %*%Matrix::Diagonal(x = 1/std_cheap)
nonstat_est_sigma <- sqrt(Matrix::diag(nonstat_cov))

sum(abs(std_cheap - std_expensive))
plot(std_cheap - std_expensive)
```


```{r}
graph_pcr_1$build_mesh(h = 0.1)
graph_pcr_2$build_mesh(h = 0.1)
pcr_1 <- compute_pcr_2v(cov_matrix = Sigma, graph = graph_pcr_1, cor_threshold = 0.1)
pcr_2 <- compute_pcr_2v(cov_matrix = nonstat_cov, graph = graph_pcr_2, cor_threshold = 0.1)
sum(abs(pcr_1 - pcr_2))
plot(pcr_1 - pcr_2)
```


The following are the plots for the marginal standard deviation.

:::: {style="display: grid; grid-template-columns: 400px 400px 400px; grid-column-gap: 1px;"}


::: {}

```{r, out.width = "100%"}
a = 2.35
graph$plot_function(X = std_expensive, vertex_size = 1, type = "plotly", line_color = "darkgreen", line_width = 3, edge_color = "black", edge_width = 3) %>% # true one
  graph$plot_function(X = std_cheap, p = ., vertex_size = 1, type = "plotly", line_color = "blue", line_width = 3, edge_color = "black", edge_width = 3) %>% # approx one
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\sigma(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```


:::


::: {}


```{r, out.width = "100%"}
graph$plot_function(X = nonstat_est_sigma, vertex_size = 1, type = "plotly", line_color = "darkgreen", line_width = 3, edge_color = "black", edge_width = 3) %>%  
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\sigma_u(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```

:::



::: {}

```{r, out.width = "100%"}
graph$plot_function(X = pcr_1, vertex_size = 1, type = "plotly", line_color = "red", line_width = 3, edge_color = "black", edge_width = 3) %>% # true one
  graph$plot_function(X = pcr_2, p = ., vertex_size = 1, type = "plotly", line_color = "red", line_width = 3, edge_color = "black", edge_width = 3) %>% # approx one
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\rho(s), \\rho_u(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```


:::

::::


The following are the corresponding plots for the practical correlation range.


:::: {style="display: grid; grid-template-columns: 400px 400px 400px; grid-column-gap: 1px;"}


::: {}

```{r, out.width = "100%"}
a = 2.35
graph$plot_function(X = pcr_1, vertex_size = 1, type = "plotly", line_color = "blue", line_width = 3, edge_color = "black", edge_width = 3) %>% 
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\rho(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```


:::


::: {}


```{r, out.width = "100%"}
graph$plot_function(X = pcr_2, vertex_size = 1, type = "plotly", line_color = "darkgreen", line_width = 3, edge_color = "black", edge_width = 3) %>%  
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\rho_u(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```

:::


::: {}

```{r, out.width = "100%"}
graph$plot_function(X = pcr_1 - pcr_2, vertex_size = 1, type = "plotly", line_color = "red", line_width = 3, edge_color = "black", edge_width = 3) %>%  
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\rho_u(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```


:::

::::


```{r}
Sigma <- cov2cor(Sigma)
```

```{r, out.width = "100%"}
graph$plot_function(X = Sigma[300,], vertex_size = 1, type = "plotly", line_color = "red", line_width = 3, edge_color = "black", edge_width = 3) %>%
  graph$plot_function(X = Sigma[5,], p = ., vertex_size = 1, type = "plotly", line_color = "darkgreen", line_width = 3, edge_color = "black", edge_width = 3) %>% 
graph$plot_function(X = Sigma[3,], p = ., vertex_size = 1, type = "plotly", line_color = "blue", line_width = 3, edge_color = "black", edge_width = 3) %>% 
  config(mathjax = 'cdn') %>% 
  layout(title = TeX("\\rho_u(s)"), 
font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           aspectratio = list(x = 1, y = 1, z = 1),
camera = list(
      eye = list(x = -0.7*a, y = 0.7*a, z = 0.7*a))))
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

