---
title: "Non-stationarity"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Go back to the [About page](about.html).

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  #out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)

library(here)
library(rmarkdown)
library(grateful) # Cite all loaded packages

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```

# Build the graph

```{r}
edges <- logo_lines() # This is a function from MetricGraph package that returns a list of edges
logo_graph <- metric_graph$new(edges = edges) # Create a new graph object

#logo_graph$prune_vertices()
logo_graph$build_mesh(h = 0.05) # Build the mesh

# Extract the mesh locations in Euclidean coordinates
xypoints <- logo_graph$mesh$V

# Plot graph with mesh
logo_graph$plot(mesh = TRUE) + 
  ggtitle("Mesh") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

# Simulate a Whittle–Matérn field

Below we simulate a Whittle–Matérn field $\dot{u}$ (`cov`) from the model

\begin{equation}
\label{SPDE}\tag{1}
(\kappa^2-\Delta)^{\alpha/2}\tau \dot{u} = \mathcal{W}.
\end{equation}

We use the `matern` parameterization, where the parameters are $\nu$ (`nu`), $\sigma$ (`sigma`), and $\rho$ (`range`), and are related to the `spde` parameterization \eqref{SPDE} as follows.

- $\alpha = \nu + 1/2$,
- $\rho = \dfrac{\sqrt{8\nu}}{\kappa}$,
- $\tau^2 = \dfrac{\Gamma(\nu)}{\sigma^2 \kappa^{2\nu}(4\pi)^{d/2}\Gamma(\nu + d/2)}$.

```{r}
# Set the parameters
sigma <- 0.1 # Standard deviation
range <- 10 # Range
nu <- 1.5 # Smoothness parameter
theta <- c(1, 1, -0.5, 0.5) # Log-regression coefficients
nu_ns <- 1.5 # Smoothness parameter for non-stationarity
cor_threshold <- 0.1 # Define correlation threshold
sigma.e <- 0.1 # Standard deviation of the error term
```

```{r}
op <- matern.operators(nu = 0.5, 
                       sigma = sigma,
                       range = range,  
                       parameterization = "matern",
                       graph = logo_graph)  

est_cov_matrix <- op$covariance_mesh()
# Precision matrix
Q <- precision(op)
rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ]
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
aux_covf1 <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 1, line_color = "blue", line_width = 2)

op <- matern.operators(nu = 2.5, 
                       sigma = sigma,
                       range = range,  
                       parameterization = "matern",
                       graph = logo_graph)  

est_cov_matrix <- op$covariance_mesh()
# Precision matrix
Q <- precision(op)
rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ]
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
aux_covf2 <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 1, line_color = "red", line_width = 2, p = aux_covf1)
```



```{r}
# Construct the approximation
op <- matern.operators(nu = nu, 
                       sigma = sigma,
                       range = range,  
                       parameterization = "matern",
                       graph = logo_graph)    

# Simulate the field
cov <- as.vector(simulate(op))

est_cov_matrix <- op$covariance_mesh()
# Precision matrix
Q <- precision(op)
rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ]
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
```


```{r}
# Plot the field
logo_graph$plot_function(X = cov, vertex_size = 0, plotly = FALSE) + 
  ggtitle("Simulated field") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

```{r}
cfo <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 1, line_color = "darkgreen", line_width = 2, p = aux_covf2)
cfo
```

# Build a non-stationary covariance function for a generalized Whittle–Matérn field

We use the simulated field to build a non-stationary covariance function for a generalized Whittle–Matérn field $u$. That is, we consider the model 

\begin{equation}
\label{NS-SPDE}\tag{2}
(\kappa^2(s)-\Delta)^{\alpha/2}\tau(s) u = \mathcal{W},
\end{equation}

with non-stationary parameters (in the `matern` parameterization) given by

\begin{equation}
\label{logregressions}
    \begin{aligned}
    \log(\sigma(s)) &= \theta_1 + \theta_3 \text{cov}(s),\\
    \log(\rho(s)) &= \theta_2 + \theta_4 \text{cov}(s).
\end{aligned}
\end{equation}

Take into account that `theta` $= (\theta_1, \theta_2, \theta_3, \theta_4)$.

```{r}
# Non-stationary parameters
B.sigma = cbind(0, 1, 0, cov, 0)
B.range = cbind(0, 0, 1, 0, cov)
```


```{r}
# Construct the approximation 
ns_op <- rSPDE::spde.matern.operators(graph = logo_graph,
                                                parameterization = "matern",
                                                B.sigma = B.sigma,
                                                B.range = B.range,
                                                theta = theta, 
                                                #m = 20, type_rational_approximation = "brasil",
                                                nu = 0.5)
est_cov_matrix <- ns_op$covariance_mesh()
# Precision matrix
Q <- precision(ns_op)
rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ]
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
aux_covf1 <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 3, line_color = "blue", line_width = 3)

# Construct the approximation 
ns_op <- rSPDE::spde.matern.operators(graph = logo_graph,
                                                parameterization = "matern",
                                                B.sigma = B.sigma,
                                                B.range = B.range,
                                                theta = theta,
                                                #m = 20, type_rational_approximation = "brasil",
                                                nu = 2.5)
est_cov_matrix <- ns_op$covariance_mesh()
# Precision matrix
Q <- precision(ns_op)
rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ]
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
aux_covf2 <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 3, line_color = "red", line_width = 3, p = aux_covf1)
```


```{r}
# Construct the approximation 
ns_op <- rSPDE::spde.matern.operators(graph = logo_graph,
                                                parameterization = "matern",
                                                B.sigma = B.sigma,
                                                B.range = B.range,
                                                theta = theta,
                                                #m = 20, type_rational_approximation = "brasil",
                                                nu = nu_ns)

# Covariance matrix
est_cov_matrix <- ns_op$covariance_mesh()

# Precision matrix
Q <- precision(ns_op)

rownumber = 289 # Choose a row number, any berween 1 and dim(est_cov_matrix)[1]
rowfromcov = est_cov_matrix[rownumber, ] # Get a row from the covariance matrix
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
sum(rowfromcov - rowfromQ) # Check they are the same
```


## Plot the covariance between a point and all other points

```{r}
# Plot the covariance between a point and all other points
logo_graph$plot_function(X = rowfromQ, vertex_size = 0) + 
  ggtitle("Covariance between a point and all other points") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

```{r}
cfo <- logo_graph$plot_function(X = rowfromcov, vertex_size = 1, plotly = TRUE, edge_color = "black", edge_width = 3, line_color = "darkgreen", line_width = 3, p = aux_covf2) %>% 
  config(mathjax = 'cdn') %>% 
  layout(font = list(family = "Palatino"),
         showlegend = FALSE,
         scene = list(
           annotations = list(
             list(
               x = 3, y = 4, z = 8,
               text = TeX("\\nu"),
               textangle = 0, ax = 60, ay = 0,
               font = list(color = "black", size = 16),
               arrowcolor = "white", arrowsize = 1, arrowwidth = 0.1,arrowhead = 1),
             list(
               x = 3, y = 4, z = 6.5,
               text = TeX("\\bullet\\mbox{ 0.5}"),
               textangle = 0, ax = 60, ay = 0,
               font = list(color = "blue", size = 16),
               arrowcolor = "white", arrowsize = 1, arrowwidth = 0.1,arrowhead = 1),
             list(
               x = 3, y = 4, z = 7,
               text = TeX("\\bullet\\mbox{ 1.5}"),
               textangle = 0, ax = 60, ay = 0,
               font = list(color = "darkgreen", size = 16),
               arrowcolor = "white", arrowsize = 1, arrowwidth = 0.1,arrowhead = 1),
             list(
               x = 3, y = 4, z = 7.5,
               text = TeX("\\bullet\\mbox{ 2.5}"),
               textangle = 0, ax = 60, ay = 0,
               font = list(color = "red", size = 16),
               arrowcolor = "white", arrowsize = 1, arrowwidth = 0.1,arrowhead = 1),
             list(
               x = 3, y = 4, z = 0,
               text = TeX("s_0"),
               textangle = 0, ax = 0, ay = 15,
               font = list(color = "black", size = 16),
               arrowcolor = "white", arrowsize = 1, arrowwidth = 0.1,arrowhead = 1)))) %>% 
  add_trace(x = xypoints[rownumber, 2], y = xypoints[rownumber, 1], z = 0, mode = "markers", type = "scatter3d", 
            marker = list(size = 4, color = "black", symbol = 104))
```

```{r}
cfo
```

## Compute the practical correlation range and the standard deviation

```{r}
# Build the models for sigma and range
model_for_sigma <- exp(B.sigma[,-1]%*%theta)
model_for_range <- exp(B.range[,-1]%*%theta)

# Compute distance matrix on the mesh
logo_graph$compute_geodist_mesh()

# Get the distance matrix
dist_matrix <- logo_graph$mesh$geo_dist

# Initialize vector to store the practical correlation range
est_range = rep(NA, dim(est_cov_matrix)[1])
  
# Compute the practical correlation range
for (i in 1:dim(est_cov_matrix)[1]) {
  est_range[i] <- data.frame(cor = est_cov_matrix[i,], dist = dist_matrix[i,]) %>% 
    arrange(dist) %>% 
    filter(cor < cor_threshold) %>%
  slice(1) %>%
  pull(dist)
}

# Compute the standard deviation
est_sigma <- sqrt(diag(INLA::inla.qinv(Q))) # It can also be computed as sqrt(Matrix::diag(est_cov_matrix))
```


## Plot the practical correlation range and the standard deviation

```{r, fig.dim = c(9,4)}
# Extract the mesh locations in Euclidean coordinates
xypoints <- logo_graph$mesh$V

# Define data frames for plotting
sd_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_sigma, .group = 1)
rho_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_range, .group = 2)
data_param_cov <- rbind(sd_df, rho_df)

group_labels <- c("Marginal standard deviation", "Practical correlation range")

min_x <- min(data_param_cov$x) |> round(2)
max_x <- max(data_param_cov$x) |> round(2)
min_y <- min(data_param_cov$y) |> round(2)
max_y <- max(data_param_cov$y) |> round(2)

# Create a base plot
base_plot <- ggplot() +
  geom_point(data = sd_df, aes(x = x, y = y), color = "black", size = 0.5, shape = 1) +
  scale_x_continuous(labels = function(x) paste0(x), name = "x", breaks = seq(min_x, max_x, by = 2)) +  
  scale_y_continuous(labels = function(y) paste0(y), name = "y", breaks = seq(min_y, max_y, by = 2)) +
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

# Create individual plots for each group and overlay data points
group_plots <- lapply(unique(data_param_cov$.group), function(gr) {
  geom_point(data = subset(data_param_cov, .group == gr), aes(x = x, y = y, color = z), size = 1)
})

# Combine the base plot and the group plots
combined_plot <- base_plot + 
  facet_wrap(~ .group, ncol = 2, labeller = labeller(.group = function(variable, value) {
    return(group_labels[value])
  })) +  # Specify custom labels for facet_wrap
  group_plots +
  scale_color_gradient(name = "") +
  labs(color = "") +
  scale_color_viridis_c(option = "D")

# Print the combined plot
print(combined_plot)
```


```{r}
ggsave(here("data_files/stand_and_range.png"), plot = combined_plot, dpi = 300)
```


## Plot tau and kappa

```{r, fig.dim = c(9,4)}
# Extract the mesh locations in Euclidean coordinates
xypoints <- logo_graph$mesh$V

est_kappa <- sqrt(8*nu_ns)/est_range
est_tau <- sqrt(gamma(nu_ns) / (est_sigma^2 * est_kappa^(2*nu_ns) * (4*pi)^(1/2) * gamma(nu_ns + 1/2)))

# Define data frames for plotting
sd_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_tau, .group = 1)
rho_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_kappa, .group = 2)
data_param_cov <- rbind(sd_df, rho_df)

group_labels <- c("Tau", "Kappa")

min_x <- min(data_param_cov$x) |> round(2)
max_x <- max(data_param_cov$x) |> round(2)
min_y <- min(data_param_cov$y) |> round(2)
max_y <- max(data_param_cov$y) |> round(2)

# Create a base plot
base_plot <- ggplot() +
  geom_point(data = sd_df, aes(x = x, y = y), color = "black", size = 0.5, shape = 1) +
  scale_x_continuous(labels = function(x) paste0(x), name = "x", breaks = seq(min_x, max_x, by = 2)) +  
  scale_y_continuous(labels = function(y) paste0(y), name = "y", breaks = seq(min_y, max_y, by = 2)) +
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

# Create individual plots for each group and overlay data points
group_plots <- lapply(unique(data_param_cov$.group), function(gr) {
  geom_point(data = subset(data_param_cov, .group == gr), aes(x = x, y = y, color = z), size = 1)
})

# Combine the base plot and the group plots
combined_plot <- base_plot + 
  facet_wrap(~ .group, ncol = 2, labeller = labeller(.group = function(variable, value) {
    return(group_labels[value])
  })) +  # Specify custom labels for facet_wrap
  group_plots +
  scale_color_gradient(name = "") +
  labs(color = "Value") +
  scale_color_viridis_c(option = "D")

# Print the combined plot
print(combined_plot)
```

```{r}
# Define data frames for plotting
sigma_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = model_for_sigma, .group = 1)
range_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = model_for_range, .group = 2)
sd_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_sigma, .group = 3)
rho_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_range, .group = 4)
data_param_cov <- rbind(sigma_df, range_df, sd_df, rho_df)

group_labels <- c("Model for sigma", "Model for range", "Standard deviation", "Practical correlation range")

min_x <- min(data_param_cov$x) |> round(2)
max_x <- max(data_param_cov$x) |> round(2)
min_y <- min(data_param_cov$y) |> round(2)
max_y <- max(data_param_cov$y) |> round(2)

# Create a base plot
base_plot <- ggplot() +
  geom_point(data = sigma_df, aes(x = x, y = y), color = "black", size = 0.5, shape = 1) +
  scale_x_continuous(labels = function(x) paste0(x), name = "x", breaks = seq(min_x, max_x, by = 2)) +  
  scale_y_continuous(labels = function(y) paste0(y), name = "y", breaks = seq(min_y, max_y, by = 2)) +
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

# Create individual plots for each group and overlay data points
group_plots <- lapply(unique(data_param_cov$.group), function(gr) {
  geom_point(data = subset(data_param_cov, .group == gr), aes(x = x, y = y, color = z), size = 1)
})

# Combine the base plot and the group plots
combined_plot <- base_plot + 
  facet_wrap(~ .group, ncol = 2, labeller = labeller(.group = function(variable, value) {
    return(group_labels[value])
  })) +  # Specify custom labels for facet_wrap
  group_plots +
  scale_color_gradient(name = "") +
  labs(color = "") +
  scale_color_viridis_c(option = "D")

# Print the combined plot
print(combined_plot)
```

## Simulation

## Simulate and plot the non-stationary field

```{r}
# Simulate the non-stationary field
u_non_stat <- simulate(ns_op)

# Plot the graph
pl <- logo_graph$plot(vertex_size = 1, vertex_color = "blue") + 
  ggtitle(latex2exp::TeX("MetricGraph package's logo")) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))


# Plot the non-stationary field
nsp <- logo_graph$plot_function(X = u_non_stat, vertex_size = 0) + 
  ggtitle("Simulated non-stationary process") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

plt_aux <- pl + nsp
```


```{r}
ggsave(here("data_files/graph_and_sim.png"), plot = plt_aux, dpi = 300)
```


```{r}
xy_points <- logo_graph$mesh$V
df_just_to_plot <- data.frame(x = xy_points[,1], y = xy_points[,2], z = as.vector(u_non_stat))
pbase <- ggplot() +
  geom_point(data = df_just_to_plot, aes(x = x, y = y, color = z), size = 1) +
  scale_color_gradient(name = "") +
  labs(color = "") +
  scale_color_viridis_c(option = "D") +
  theme_minimal() +
  theme(text = element_text(family = "Palatino")) +
  ggtitle("Simulated non-stationary process")

ppmss <- pl + pbase
```

```{r}
ggsave(here("data_files/graph_and_sim_p.png"), plot = ppmss, dpi = 300)
```

## Simulate observations

```{r}
# Number of observations per edge
obs.per.edge <- 50
obs.loc <- NULL
for(i in 1:logo_graph$nE) {
  obs.loc <- rbind(obs.loc,
                   cbind(rep(i,obs.per.edge), runif(obs.per.edge)))
}
n.obs <- obs.per.edge*logo_graph$nE

# Compute the FEM basis functions on the observation locations
A <- logo_graph$fem_basis(obs.loc)

# Define two covariates
x1 <- sin(obs.loc[,1])
x2 <- sin(obs.loc[,2])

# Simulate observations
Y <- 1 + 2*x1 - 3*x2 + as.vector(A %*% u_non_stat + sigma.e * rnorm(n.obs))

# Create a data frame with the observations
df_data <- data.frame(y = Y, edge_number = obs.loc[,1],
                        distance_on_edge = obs.loc[,2],
                        x1 = x1, x2 = x2)
```

## Compare the data before and after adding the observations

```{r}
# Add the observations to the graph
logo_graph$add_observations(data = df_data, normalized = TRUE, clear_obs = TRUE)
data_after_added <- logo_graph$get_data() %>% mutate(something = 1)
# Add the observations to the graph
logo_graph$add_observations(data = data_after_added, normalized = TRUE, clear_obs = TRUE)
data_after_added2 <- logo_graph$get_data()
# Get edge_number and distance_on_edge to compare
toy1 <- df_data %>% dplyr::select(edge_number, distance_on_edge)
toy2 <- data_after_added %>% rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) %>% as.data.frame() %>% dplyr::select(edge_number, distance_on_edge)
toy3 <- data_after_added2 %>% rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) %>% as.data.frame() %>% dplyr::select(edge_number, distance_on_edge)
# Compare them
identical(toy1, toy2)
identical(toy2, toy3)
```

## Add and plot the simulated data

```{r}
# Add the observations to the graph
logo_graph$add_observations(data = df_data, normalized = TRUE, clear_obs = TRUE)

# Get the data
data <- logo_graph$get_data()

# Plot the simulated observations
logo_graph$plot(data = "y", vertex_size = 0) + 
  ggtitle("Simulated data") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```


## Compute the distance matrix and create the groups

```{r}
# Define aux data frame to compute the distance matrix
aux <- data |>
  rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) |> # Rename the variables (because sf_graph$compute_geodist_PtE() requires so)
  as.data.frame() |> # Transform to a data frame (i.e., remove the metric_graph class)
  dplyr::select(edge_number, distance_on_edge)

# Compute the distance matrix
distmatrix <- logo_graph$compute_geodist_PtE(PtE = aux,
                                             normalized = TRUE,
                                             include_vertices = FALSE)
# Define the distance vector
distance = seq(from = 0, to = 2.5, by = 0.25)

GROUPS <- list()
for (j in 1:length(distance)) {
  GROUPS[[j]] = list()
  for (i in 1:nrow(aux)) {
    GROUPS[[j]][[i]] <- which(as.vector(distmatrix[i, ]) <= distance[j])
  }
}
```

## Check groups

```{r}
indexofinterest <- 2000 # Any number between 1 and nrow(data)
pointofinterest <- GROUPS[[3]][[indexofinterest]]
pointofinterest2 <- GROUPS[[11]][[indexofinterest]]
ggplot() +
  geom_point(data = data, aes(x = .coord_x, y = .coord_y), color = "black") +
  geom_point(data = data[pointofinterest2, ], aes(x = .coord_x, y = .coord_y), color = "green") +
  geom_point(data = data[pointofinterest, ], aes(x = .coord_x, y = .coord_y), color = "blue") +
  geom_point(data = data[indexofinterest, ], aes(x = .coord_x, y = .coord_y), color = "red") +
  ggtitle("Groups") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino")) +
  coord_fixed()
```


<!-- :::: {style="display: flex; to display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"} -->

<!-- ::: {} -->

## Fit the stationary model

```{r}
rspde_model_stat <- rspde.metric_graph(logo_graph,
                                       parameterization = "matern",
                                       nu = nu_ns)

data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        loc_name = "loc")
cmp_stat <- y ~ -1 +
  Intercept(1) +
  x1 +
  x2 +
  field(loc, model = rspde_model_stat)

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_stat)
summary(rspde.result(rspde_fit_stat, "field", rspde_model_stat))
```


<!-- ::: -->

<!-- ::: {} -->


## Fit the non-stationary model

```{r}
rspde_model_nonstat <- rspde.metric_graph(logo_graph,
                                          B.sigma = B.sigma,
                                          B.range = B.range,
                                          parameterization = "matern",
                                          nu = nu_ns)

data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                           loc_name = "loc")
cmp_nonstat <- y ~ -1 +
  Intercept(1) +
  x1 +
  x2 +
  field(loc, model = rspde_model_nonstat)

rspde_fit_nonstat <-
  bru(cmp_nonstat,
      data = data_rspde_bru_nonstat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))
```


<!-- ::: -->

<!-- :::: -->


## Cross-validation

```{r}
mse.stat <- mse.nonstat <- ls.stat <- ls.nonstat <- rep(0,length(distance))
# cross-validation for-loop
for (j in 1:length(distance)) {
  # cross-validation of the stationary model
  cv.stat <- inla.group.cv(rspde_fit_stat, groups = GROUPS[[j]])
  # cross-validation of the nonstationary model
  cv.nonstat <- inla.group.cv(rspde_fit_nonstat, groups = GROUPS[[j]])
  # obtain MSE and LS
  mse.stat[j] <- mean((cv.stat$mean - data$y)^2)
  mse.nonstat[j] <- mean((cv.nonstat$mean - data$y)^2)
  ls.stat[j] <- mean(log(cv.stat$cv))
  ls.nonstat[j] <- mean(log(cv.nonstat$cv))
}
```

## Plot cross-validation results

```{r}
# Create data frames
mse_df <- data.frame(
  distance,
  Statnu0.5 = mse.stat,
  Nonstatnu0.5 = mse.nonstat
)

ls_df <- data.frame(
  distance,
  Statnu0.5 = -ls.stat,
  Nonstatnu0.5 = -ls.nonstat
)

# Convert to long format
mse_long <- mse_df %>%
  pivot_longer(cols = -distance, names_to = "nu", values_to = "MSE")

ls_long <- ls_df %>%
  pivot_longer(cols = -distance, names_to = "nu", values_to = "LogScore")


# Update the label mappings with the new legend title
label_mapping <- c(
  "Statnu0.5" = "0.5", 
  "Nonstatnu0.5" = "0.5"
)

# Define color and linetype mapping
color_mapping <- c(
  "Statnu0.5" = "blue", 
  "Nonstatnu0.5" = "red"
)

linetype_mapping <- c(
  "Statnu0.5" = "dotdash", 
  "Nonstatnu0.5" = "solid"
)

# Plot MSE
mse_plot <- ggplot(mse_long, aes(x = distance, y = MSE, color = nu, linetype = nu)) +
  geom_line(linewidth = 1) +
  labs(y = "MSE", x = "Distance") +
  scale_color_manual(values = color_mapping, labels = label_mapping, name = expression(nu)) +
  scale_linetype_manual(values = linetype_mapping, labels = label_mapping, name = expression(nu)) +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))

# Plot negative log-score
ls_plot <- ggplot(ls_long, aes(x = distance, y = LogScore, color = nu, linetype = nu)) +
  geom_line(linewidth = 1) +
  labs(y = "Negative Log-Score", x = "Distance") +
  scale_color_manual(values = color_mapping, labels = label_mapping, name = expression(nu)) +
  scale_linetype_manual(values = linetype_mapping, labels = label_mapping, name = expression(nu)) +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))

# Combine plots with a shared legend at the top in a single line
combined_plot <- mse_plot + ls_plot + 
  plot_layout(guides = 'collect') & 
  theme(legend.position = 'right') & 
  guides(color = guide_legend(nrow = 2), linetype = guide_legend(nrow = 1))

# Display combined plot
print(combined_plot)
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
