---
title: "Non-stationarity"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Go back to the [About page](about.html).

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(dplyr)
library(plotly)

library(here)
library(rmarkdown)
library(grateful) # Cite all loaded packages

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```

# Build the graph

```{r, fig.height = 6}
edges <- logo_lines() # This is a function from MetricGraph package that returns a list of edges
logo_graph <- metric_graph$new(edges = edges) # Create a new graph object

#logo_graph$prune_vertices()
logo_graph$build_mesh(h = 0.05) # Build the mesh

# Plot graph with mesh
logo_graph$plot(mesh = TRUE) + 
  ggtitle("Mesh") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

# Simulate a Whittle–Matérn field

Below we simulate a Whittle–Matérn field $\dot{u}$ (`cov`) from the model

\begin{equation}
\label{SPDE}\tag{1}
(\kappa^2-\Delta)^{\alpha/2}\tau \dot{u} = \mathcal{W}.
\end{equation}

We use the `matern` parameterization, where the parameters are $\nu$ (`nu`), $\sigma$ (`sigma`), and $\rho$ (`range`), and are related to the `spde` parameterization \eqref{SPDE} as follows.

- $\alpha = \nu + 1/2$,
- $\rho = \dfrac{\sqrt{8\nu}}{\kappa}$,
- $\tau^2 = \dfrac{\Gamma(\nu)}{\sigma^2 \kappa^{2\nu}(4\pi)^{d/2}\Gamma(\nu + d/2)}$.

```{r}
sigma <- 0.1 
range <- 10 
nu <- 1.5 # Smoothness parameter

# Construct the approximation
op <- matern.operators(nu = nu, 
                       sigma = sigma,
                       range = range,  
                       parameterization = "matern",
                       graph = logo_graph)    

# Simulate the field
cov <- as.vector(simulate(op))

# Plot the field
logo_graph$plot_function(X = cov, vertex_size = 0, plotly = F) + 
  ggtitle("Simulated field") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

# Build a non-stationary covariance function for a generalized Whittle–Matérn field

We use the simulated field to build a non-stationary covariance function for a generalized Whittle–Matérn field $u$. That is, we consider the model 

\begin{equation}
\label{NS-SPDE}\tag{2}
(\kappa^2(s)-\Delta)^{\alpha/2}\tau(s) u = \mathcal{W},
\end{equation}

with non-stationary parameters (in the `matern` parameterization) given by

\begin{equation}
\label{logregressions}
    \begin{aligned}
    \log(\sigma(s)) &= \theta_1 + \theta_3 \text{cov}(s),\\
    \log(\rho(s)) &= \theta_2 + \theta_4 \text{cov}(s).
\end{aligned}
\end{equation}

Below `theta` $= (\theta_1, \theta_2, \theta_3, \theta_4)$.

```{r, fig.height = 4}
# Non-stationary parameters
B.sigma = cbind(0, 1, 0, cov, 0)
B.range = cbind(0, 0, 1, 0, cov)

# Log-regression coefficients
theta = c(1,1,-0.5,0.5)

# Build the models for sigma and range
model_for_sigma <- exp(B.sigma[,-1]%*%theta)
model_for_range <- exp(B.range[,-1]%*%theta)
```


## Plot the models for $\sigma$ and $\rho$

```{r, fig.height = 4}
# Extract the mesh locations in Euclidean coordinates
xypoints <- logo_graph$mesh$V

# Define data frames for plotting
sigma_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = model_for_sigma, .group = 1)
range_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = model_for_range, .group = 2)
datarep <- rbind(sigma_df, range_df)

# Plot details
group_labels <- c("Model for sigma", "Model for range")

# Plot details
min_x <- min(datarep$x) |> round(2)
max_x <- max(datarep$x) |> round(2)
min_y <- min(datarep$y) |> round(2)
max_y <- max(datarep$y) |> round(2)

# Create a base plot 
base_plot <- ggplot() +
  geom_point(data = sigma_df, aes(x = x, y = y), color = "black", size = 0.5, shape = 1) +
  scale_x_continuous(labels = function(x) paste0(x), name = "x", breaks = seq(min_x, max_x, by = 2)) +  
  scale_y_continuous(labels = function(y) paste0(y), name = "y", breaks = seq(min_y, max_y, by = 2)) +
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

# Create individual plots for each group and overlay data points
group_plots <- lapply(unique(datarep$.group), function(gr) {
  geom_point(data = subset(datarep, .group == gr), aes(x = x, y = y, color = z), size = 1)  
})

# Combine the base plot and the group plots
combined_plot <- base_plot + 
  facet_wrap(~ .group, ncol = 2, labeller = labeller(.group = function(variable, value) {
    return(group_labels[value])
  })) + 
  group_plots +
  scale_color_gradient(name = "") +
  labs(color = "") +
  scale_color_viridis_c(option = "D")

# Print the combined plot
print(combined_plot)
```

# Non-stationary covariance

```{r}
nu <- 0.5
# Construct the approximation 
ns_op <- rSPDE::spde.matern.operators(graph = logo_graph,
                                                parameterization = "matern",
                                                B.sigma = B.sigma,
                                                B.range = B.range,
                                                theta = theta,
                                                nu = nu)

# Covariance matrix
est_cov_matrix <- ns_op$covariance_mesh()

# Precision matrix
Q <- precision(ns_op)
```

## Plot the covariance between a point and all other points

```{r}
rownumber = 300 # Choose a row number
rowfromcov = est_cov_matrix[rownumber, ] # Get a row from the covariance matrix
rowfromQ = solve(Q, replace(numeric(dim(Q)[1]), rownumber, 1)) # Another way to get a row from the covariance matrix
sum(rowfromcov - rowfromQ) # Check they are the same
```

```{r}
# Plot the covariance between a point and all other points
logo_graph$plot_function(X = rowfromQ, vertex_size = 0) + 
  ggtitle("Covariance between a point and all other points") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

## Compute the practical correlation range and the standard deviation

```{r}
# Compute distance matrix on the mesh
logo_graph$compute_geodist_mesh()

# Get the distance matrix
dist_matrix <- logo_graph$mesh$geo_dist

# Define correlation threshold
cor_threshold <- 1

# Initialize vector to store the practical correlation range
est_range = rep(NA, dim(est_cov_matrix)[1])
  
# Compute the practical correlation range
for (i in 1:dim(est_cov_matrix)[1]) {
  rowi <- est_cov_matrix[i,]
  sorted <- sort(rowi, decreasing = TRUE, index.return = TRUE)
  index_cor_smaller_pcr <- which(sorted$x < cor_threshold)[1]
  pos_dist_matrix <- sorted$ix[index_cor_smaller_pcr]
  est_range[i] <- dist_matrix[i, pos_dist_matrix]
}

# Compute the standard deviation
est_sigma <- sqrt(diag(INLA::inla.qinv(Q))) # It can also be computed as sqrt(Matrix::diag(est_cov_matrix))
```

## Plot the practical correlation range and the standard deviation

```{r, fig.height = 4}
sd_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_sigma, .group = 1)
rho_df <- data.frame(x = xypoints[,1], y = xypoints[,2], z = est_range, .group = 2)


datarep <- rbind(sd_df, rho_df)
group_labels <- c("Standard deviation", "Practical correlation range")

min_x <- min(datarep$x) |> round(2)
max_x <- max(datarep$x) |> round(2)
min_y <- min(datarep$y) |> round(2)
max_y <- max(datarep$y) |> round(2)

# Create a base plot
base_plot <- ggplot() +
  geom_point(data = sd_df, aes(x = x, y = y), color = "black", size = 0.5, shape = 1) +
  scale_x_continuous(labels = function(x) paste0(x), name = "x", breaks = seq(min_x, max_x, by = 2)) +  
  scale_y_continuous(labels = function(y) paste0(y), name = "y", breaks = seq(min_y, max_y, by = 2)) +
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))

# Create individual plots for each group and overlay data points
group_plots <- lapply(unique(datarep$.group), function(gr) {
  geom_point(data = subset(datarep, .group == gr), aes(x = x, y = y, color = z), size = 1)
})

# Combine the base plot and the group plots
combined_plot <- base_plot + 
  facet_wrap(~ .group, ncol = 2, labeller = labeller(.group = function(variable, value) {
    return(group_labels[value])
  })) +  # Specify custom labels for facet_wrap
  group_plots +
  scale_color_gradient(name = "") +
  labs(color = "") +
  scale_color_viridis_c(option = "D")

# Print the combined plot
print(combined_plot)
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
