---
title: "Numerical error"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Go back to the [About page](about.html). This [link](../data_files/README.html) might be useful to keep track of the files created during the preprocessing.

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "120%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
library(MetricGraph)
library(Matrix)
library(rSPDE)

library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(latex2exp)
library(ggtext)

library(grateful) # Cite all loaded packages
library(here) # here() starts from the home directory

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```


# Interval graph


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
gets_graph_interval <- function(n){
  edge <- rbind(c(0,0),c(1,0))
  edges = list(edge)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(n = n)
  return(graph)
}

# matern covariance function. Same as in the package
matern.covariance <- function(h, kappa, nu, sigma) {
  if (nu == 1 / 2) {
    C <- sigma^2 * exp(-kappa * abs(h))
  } else {
    C <- (sigma^2 / (2^(nu - 1) * gamma(nu))) *
      ((kappa * abs(h))^nu) * besselK(kappa * abs(h), nu)
  }
  C[h == 0] <- sigma^2
  return(as.matrix(C))
}

# folded.matern.covariance.1d I edited
folded.matern.covariance.1d.local <- function(x, kappa, nu, sigma,
                                              L = 1, N = 10,
                                              boundary = c("neumann",
                                                           "dirichlet", "periodic")) {
  boundary <- tolower(boundary[1])
  if (!(boundary %in% c("neumann", "dirichlet", "periodic"))) {
    stop("The possible boundary conditions are 'neumann',
    'dirichlet' or 'periodic'!")
  }
  addi = t(outer(x, x, "+"))
  diff = t(outer(x, x, "-"))
  s1 <- sapply(-N:N, function(j) { 
    diff + 2 * j * L
  })
  s2 <- sapply(-N:N, function(j) {
    addi + 2 * j * L
  })
  if (boundary == "neumann") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) +
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else if (boundary == "dirichlet") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) -
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else {
    C <- rowSums(matern.covariance(h = s1,
                                   kappa = kappa, nu = nu, sigma = sigma))
  }
  return(matrix(C, nrow = length(x)))
}

# function 2
gets_true_cov_mat = function(graph, kappa, nu, sigma, N, boundary){
  h = graph$mesh$V[,1] 
  true_cov_mat = folded.matern.covariance.1d.local(x = h, kappa = kappa, nu = nu, sigma = sigma, N = N, boundary = boundary)
  return(true_cov_mat)
}


# parameters
n_vector = c(998, 198, 98, 8)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05)
boundary = "neumann"
sigma = 1
N.folded = 10

Error = list()
for (s in c(1,2,3,4)) { # loop over n_vector
  Error[[as.character(n_vector[s])]] = list()
  n = n_vector[s]
  graph = gets_graph_interval(n = n)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(n_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(1,2,3,4)) { # loop over rho_vector
        Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (r in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[r]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
              # getting true covariance
              true_cov_mat = gets_true_cov_mat(graph = graph,
                                               kappa = kappa,
                                               nu = nu,
                                               sigma = sigma,
                                               N = N.folded,
                                               boundary = boundary)
              
              # getting the approximate covariance matrix
              op = matern.operators(alpha = alpha, 
                                    kappa = kappa, 
                                    tau = tau,
                                    m = m, 
                                    graph = graph,
                                    type = type,
                                    type_rational_approximation = type_rational_approximation)
              
              appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,1] = L_inf_error
            Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,1] = NA
              Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,2] = NA
            })
            print(paste("n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}

Error_interval_DEF = Error
save(Error_interval_DEF, file = here("data_files/Error_interval_DEF.RData"))
```

## Plotting the errors

```{r, class.source = "fold-hide"}
load(here("data_files/Error_interval_DEF.RData"))

n_vector = c("998", "198", "98", "8") 
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
nu_vector = seq(0.1,2.49,by=0.05)[1:29]

n = "998"
t = "chebfun"


dat = rbind(
data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,1],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

## L2

data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,2],
                     Error_interval_DEF[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
)
)


dat$rho = factor(dat$rho, levels = c("0.1", "0.5", "1", "2"))
dat$L = factor(dat$L, levels = c("L2", "Linf"))
levels(dat$rho) = c("0.1" = latex2exp::TeX("$\\rho = 0.1$"), "0.5" = latex2exp::TeX("$\\rho = 0.5$"), "1" = latex2exp::TeX("$\\rho = 1$"), "2" = latex2exp::TeX("$\\rho = 2$"))
levels(dat$L) = c("L2" = latex2exp::TeX("$L_2(\\Gamma\\times\\Gamma)$ error"), "Linf" = latex2exp::TeX("$L_\\infty(\\Gamma\\times\\Gamma)$ error"))


ggplot(dat, aes(nu, error, colour = as.factor(m))) + 
  facet_grid(L ~ rho, labeller = label_parsed) +
  geom_line() +
  scale_y_log10(n.breaks = 7) +
  scale_x_continuous(n.breaks = 6) +
  theme_bw() +
  theme(panel.spacing = unit(0.3, "cm"), text = element_text(family = "Palatino")) +
  labs(x = bquote(nu ~ "(smoothness parameter)"), y = "error", color = "Order (m)") + ggtitle("Interval graph")
```


# Circle graph


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
gets_graph_circle <- function(n){
  r = 1/(pi)
  theta <- seq(from=-pi,to=pi,length.out = 100)
  edge <- cbind(1+r+r*cos(theta),r*sin(theta))
  edges = list(edge)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(n = n)
  return(graph)
}

# matern covariance function. Same as in the package
matern.covariance <- function(h, kappa, nu, sigma) {
  if (nu == 1 / 2) {
    C <- sigma^2 * exp(-kappa * abs(h))
  } else {
    C <- (sigma^2 / (2^(nu - 1) * gamma(nu))) *
      ((kappa * abs(h))^nu) * besselK(kappa * abs(h), nu)
  }
  C[h == 0] <- sigma^2
  return(as.matrix(C))
}

# folded.matern.covariance.1d I edited
folded.matern.covariance.1d.local <- function(x, kappa, nu, sigma, L = 1, N = 10, boundary = c("neumann",
                                                                                               "dirichlet", "periodic")) {
  boundary <- tolower(boundary[1])
  if (!(boundary %in% c("neumann", "dirichlet", "periodic"))) {
    stop("The possible boundary conditions are 'neumann',
    'dirichlet' or 'periodic'!")
  }
  addi = t(outer(x, x, "+"))
  diff = t(outer(x, x, "-"))
  s1 <- sapply(-N:N, function(j) { # s1 is a matrix of size length(h)x(2N+1)
    diff + 2 * j * L
  })
  s2 <- sapply(-N:N, function(j) {
    addi + 2 * j * L
  })
  if (boundary == "neumann") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) +
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else if (boundary == "dirichlet") {
    C <- rowSums(matern.covariance(h = s1, kappa = kappa,
                                   nu = nu, sigma = sigma) -
                   matern.covariance(h = s2, kappa = kappa,
                                     nu = nu, sigma = sigma))
  } else {
    C <- rowSums(matern.covariance(h = s1,
                                   kappa = kappa, nu = nu, sigma = sigma))
  }
  return(matrix(C, nrow = length(x)))
}

# function 2
gets_true_cov_mat = function(graph, kappa, nu, sigma, N, boundary){
  h = c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[,2])
  true_cov_mat = folded.matern.covariance.1d.local(x = h, kappa = kappa, nu = nu, sigma = sigma, N = N, boundary = boundary)
  return(true_cov_mat)
}


# parameters
n_vector = 2*c(998, 198, 98, 8)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05)
boundary = "periodic"
sigma = 1
N.folded = 10

Error = list()
for (s in c(1,2,3,4)) { # loop over n_vector
  Error[[as.character(n_vector[s])]] = list()
  n = n_vector[s]
  graph = gets_graph_circle(n = n)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(n_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(1,2,3,4)) { # loop over rho_vector
        Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (r in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[r]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
              # getting true covariance
              true_cov_mat = gets_true_cov_mat(graph = graph,
                                               kappa = kappa,
                                               nu = nu,
                                               sigma = sigma,
                                               N = N.folded,
                                               boundary = boundary)
              
              # getting the approximate covariance matrix
              op = matern.operators(alpha = alpha, 
                                    kappa = kappa, 
                                    tau = tau,
                                    m = m, 
                                    graph = graph,
                                    type = type,
                                    type_rational_approximation = type_rational_approximation)
              
              appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,1] = L_inf_error
            Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,1] = NA
              Error[[as.character(n_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][r,2] = NA
            })
            print(paste("n=", n, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}
Error_circle_DEF = Error
save(Error_circle_DEF, file = here("data_files/Error_circle_DEF.RData"))
```


## Plotting the errors

```{r, class.source = "fold-hide"}
load(here("data_files/Error_circle_DEF.RData"))

n_vector = c("1996", "396", "196", "16") 
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
nu_vector = seq(0.1,2.49,by=0.05)[1:29]

n = "1996"
t = "chebfun"
  

dat = rbind(
data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,1],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

## L2

data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,2],
                     Error_circle_DEF[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
)
)


dat$rho = factor(dat$rho, levels = c("0.1", "0.5", "1", "2"))
dat$L = factor(dat$L, levels = c("L2", "Linf"))
levels(dat$rho) = c("0.1" = latex2exp::TeX("$\\rho = 0.1$"), "0.5" = latex2exp::TeX("$\\rho = 0.5$"), "1" = latex2exp::TeX("$\\rho = 1$"), "2" = latex2exp::TeX("$\\rho = 2$"))
levels(dat$L) = c("L2" = latex2exp::TeX("$L_2(\\Gamma\\times\\Gamma)$ error"), "Linf" = latex2exp::TeX("$L_\\infty(\\Gamma\\times\\Gamma)$ error"))


ggplot(dat, aes(nu, error, colour = as.factor(m))) + 
  facet_grid(L ~ rho, labeller = label_parsed) +
  geom_line() +
  scale_y_log10(n.breaks = 7) +
  scale_x_continuous(n.breaks = 6) +
  theme_bw() +
  theme(panel.spacing = unit(0.3, "cm"), text = element_text(family = "Palatino")) +
  labs(x = bquote(nu ~ "(smoothness parameter)"), y = "error", color = "Order (m)") + ggtitle("Circle graph")
```

# Tadpole graph


## `rho = 0.1`

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
tadpole.eig <- function(k,graph){
  x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
  x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 
  
  if(k==0){ 
    f.e1 <- rep(1,length(x1)) 
    f.e2 <- rep(1,length(x2)) 
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    f = list(phi=f1/sqrt(3)) 
    
  } else {
    f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
    f.e2 <- sin(pi*k*x2/2)                  
    
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    
    if((k %% 2)==1){ 
      f = list(phi=f1/sqrt(3)) 
    } else { 
      f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
      f.e2 <- cos(pi*k*x2/2)
      f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
      f <- list(phi=f1,psi=f2/sqrt(3/2))
    }
  }
  
  return(f)
}

# function 2
gets_graph <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 100)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges = list(edge1, edge2)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(h=h)
  return(graph)
}

#function 3
gets_true_cov_mat <- function(graph, kappa, tau, alpha, n.overkill){
  Sigma.kl <- matrix(0,nrow = dim(graph$mesh$V)[1],ncol = dim(graph$mesh$V)[1])
  for(i in 0:n.overkill){
    phi <- tadpole.eig(i,graph)$phi
    Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*phi%*%t(phi)
    if(i>0 && (i %% 2)==0){ 
      psi <- tadpole.eig(i,graph)$psi
      Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*psi%*%t(psi)
    }
    
  }
  Sigma.kl <- Sigma.kl/tau^2
  return(Sigma.kl)
}

# parameters
h_vector = c(0.001, 0.005, 0.01, 0.1)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05) 
n.overkill = 1000
sigma = 1

Error = list()
for (s in c(1,2,3,4)) { # loop over h_vector
  Error[[as.character(h_vector[s])]] = list()
  h = h_vector[s]
  graph = gets_graph(h = h)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(h_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(1)) { # loop over rho_vector
        Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (n in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[n]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
            # getting true covariance
            true_cov_mat = gets_true_cov_mat(graph = graph,
                                             kappa = kappa,
                                             tau = tau,
                                             alpha = alpha,
                                             n.overkill = n.overkill)
            
            # getting the approximate covariance matrix
            op = matern.operators(alpha = alpha, 
                                  kappa = kappa, 
                                  tau = tau,
                                  m = m, 
                                  graph = graph,
                                  type = type,
                                  type_rational_approximation = type_rational_approximation)
            appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = L_inf_error
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = NA
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = NA
            })
            print(paste("h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}

Error_tadpole_DEF_rho0.1 = Error
save(Error_tadpole_DEF_rho0.1, file = here("data_files/Error_tadpole_DEF_rho0.1.RData"))
```


## `rho = 0.5`


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
tadpole.eig <- function(k,graph){
  x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
  x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 
  
  if(k==0){ 
    f.e1 <- rep(1,length(x1)) 
    f.e2 <- rep(1,length(x2)) 
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    f = list(phi=f1/sqrt(3)) 
    
  } else {
    f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
    f.e2 <- sin(pi*k*x2/2)                  
    
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    
    if((k %% 2)==1){ 
      f = list(phi=f1/sqrt(3)) 
    } else { 
      f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
      f.e2 <- cos(pi*k*x2/2)
      f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
      f <- list(phi=f1,psi=f2/sqrt(3/2))
    }
  }
  
  return(f)
}

# function 2
gets_graph <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 100)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges = list(edge1, edge2)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(h=h)
  return(graph)
}

#function 3
gets_true_cov_mat <- function(graph, kappa, tau, alpha, n.overkill){
  Sigma.kl <- matrix(0,nrow = dim(graph$mesh$V)[1],ncol = dim(graph$mesh$V)[1])
  for(i in 0:n.overkill){
    phi <- tadpole.eig(i,graph)$phi
    Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*phi%*%t(phi)
    if(i>0 && (i %% 2)==0){ 
      psi <- tadpole.eig(i,graph)$psi
      Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*psi%*%t(psi)
    }
    
  }
  Sigma.kl <- Sigma.kl/tau^2
  return(Sigma.kl)
}

# parameters
h_vector = c(0.001, 0.005, 0.01, 0.1)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05) 
n.overkill = 1000
sigma = 1

Error = list()
for (s in c(1,2,3,4)) { # loop over h_vector
  Error[[as.character(h_vector[s])]] = list()
  h = h_vector[s]
  graph = gets_graph(h = h)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(h_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(2)) { # loop over rho_vector
        Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (n in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[n]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
            # getting true covariance
            true_cov_mat = gets_true_cov_mat(graph = graph,
                                             kappa = kappa,
                                             tau = tau,
                                             alpha = alpha,
                                             n.overkill = n.overkill)
            
            # getting the approximate covariance matrix
            op = matern.operators(alpha = alpha, 
                                  kappa = kappa, 
                                  tau = tau,
                                  m = m, 
                                  graph = graph,
                                  type = type,
                                  type_rational_approximation = type_rational_approximation)
            appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = L_inf_error
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = NA
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = NA
            })
            print(paste("h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}

Error_tadpole_DEF_rho0.5 = Error
save(Error_tadpole_DEF_rho0.5, file = here("data_files/Error_tadpole_DEF_rho0.5.RData"))
```


## `rho = 1`


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
tadpole.eig <- function(k,graph){
  x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
  x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 
  
  if(k==0){ 
    f.e1 <- rep(1,length(x1)) 
    f.e2 <- rep(1,length(x2)) 
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    f = list(phi=f1/sqrt(3)) 
    
  } else {
    f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
    f.e2 <- sin(pi*k*x2/2)                  
    
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    
    if((k %% 2)==1){ 
      f = list(phi=f1/sqrt(3)) 
    } else { 
      f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
      f.e2 <- cos(pi*k*x2/2)
      f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
      f <- list(phi=f1,psi=f2/sqrt(3/2))
    }
  }
  
  return(f)
}

# function 2
gets_graph <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 100)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges = list(edge1, edge2)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(h=h)
  return(graph)
}

#function 3
gets_true_cov_mat <- function(graph, kappa, tau, alpha, n.overkill){
  Sigma.kl <- matrix(0,nrow = dim(graph$mesh$V)[1],ncol = dim(graph$mesh$V)[1])
  for(i in 0:n.overkill){
    phi <- tadpole.eig(i,graph)$phi
    Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*phi%*%t(phi)
    if(i>0 && (i %% 2)==0){ 
      psi <- tadpole.eig(i,graph)$psi
      Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*psi%*%t(psi)
    }
    
  }
  Sigma.kl <- Sigma.kl/tau^2
  return(Sigma.kl)
}

# parameters
h_vector = c(0.001, 0.005, 0.01, 0.1)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05) 
n.overkill = 1000
sigma = 1

Error = list()
for (s in c(1,2,3,4)) { # loop over h_vector
  Error[[as.character(h_vector[s])]] = list()
  h = h_vector[s]
  graph = gets_graph(h = h)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(h_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(3)) { # loop over rho_vector
        Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (n in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[n]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
            # getting true covariance
            true_cov_mat = gets_true_cov_mat(graph = graph,
                                             kappa = kappa,
                                             tau = tau,
                                             alpha = alpha,
                                             n.overkill = n.overkill)
            
            # getting the approximate covariance matrix
            op = matern.operators(alpha = alpha, 
                                  kappa = kappa, 
                                  tau = tau,
                                  m = m, 
                                  graph = graph,
                                  type = type,
                                  type_rational_approximation = type_rational_approximation)
            appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = L_inf_error
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = NA
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = NA
            })
            print(paste("h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}

Error_tadpole_DEF_rho1 = Error
save(Error_tadpole_DEF_rho1, file = here("data_files/Error_tadpole_DEF_rho1.RData"))
```


## `rho = 2`

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, eval = FALSE, class.source = "fold-hide"}
# library calls
library(MetricGraph)
library(Matrix)
library(rSPDE)

# function 1
tadpole.eig <- function(k,graph){
  x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
  x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 
  
  if(k==0){ 
    f.e1 <- rep(1,length(x1)) 
    f.e2 <- rep(1,length(x2)) 
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    f = list(phi=f1/sqrt(3)) 
    
  } else {
    f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
    f.e2 <- sin(pi*k*x2/2)                  
    
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    
    if((k %% 2)==1){ 
      f = list(phi=f1/sqrt(3)) 
    } else { 
      f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
      f.e2 <- cos(pi*k*x2/2)
      f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
      f <- list(phi=f1,psi=f2/sqrt(3/2))
    }
  }
  
  return(f)
}

# function 2
gets_graph <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 100)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges = list(edge1, edge2)
  graph <- metric_graph$new(edges = edges)
  graph$build_mesh(h=h)
  return(graph)
}

#function 3
gets_true_cov_mat <- function(graph, kappa, tau, alpha, n.overkill){
  Sigma.kl <- matrix(0,nrow = dim(graph$mesh$V)[1],ncol = dim(graph$mesh$V)[1])
  for(i in 0:n.overkill){
    phi <- tadpole.eig(i,graph)$phi
    Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*phi%*%t(phi)
    if(i>0 && (i %% 2)==0){ 
      psi <- tadpole.eig(i,graph)$psi
      Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*psi%*%t(psi)
    }
    
  }
  Sigma.kl <- Sigma.kl/tau^2
  return(Sigma.kl)
}

# parameters
h_vector = c(0.001, 0.005, 0.01, 0.1)
type_vector = c("covariance", "operator")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
rho_vector = c(0.1, 0.5, 1, 2)
m_vector = c(0,1,2,3,4)
nu_vector = seq(0.1,2.49,by=0.05) 
n.overkill = 1000
sigma = 1

Error = list()
for (s in c(1,2,3,4)) { # loop over h_vector
  Error[[as.character(h_vector[s])]] = list()
  h = h_vector[s]
  graph = gets_graph(h = h)
  for (i in c(1)) { # loop over type_vector
    Error[[as.character(h_vector[s])]][[type_vector[i]]] = list()
    type = type_vector[i]
    for (j in c(1,2,3)) { # loop over type_rational_approximation_vector
      Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]] = list()
      type_rational_approximation = type_rational_approximation_vector[j]
      for (k in c(4)) { # loop over rho_vector
        Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]] = list()
        rho = rho_vector[k]
        for (l in 1:length(m_vector)) { # loop over m_vector
          Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]] = matrix(NA, nrow = length(nu_vector), ncol = 2)
          colnames(Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]]) = c("L_inf_error", "L_2_error")
          m = m_vector[l]
          for (n in 1:length(nu_vector)) { # loop over nu_vector
            nu = nu_vector[n]
            
            kappa = sqrt(8*nu)/rho
            tau = sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
            alpha = nu + 1/2
            
            tryCatch({
            # getting true covariance
            true_cov_mat = gets_true_cov_mat(graph = graph,
                                             kappa = kappa,
                                             tau = tau,
                                             alpha = alpha,
                                             n.overkill = n.overkill)
            
            # getting the approximate covariance matrix
            op = matern.operators(alpha = alpha, 
                                  kappa = kappa, 
                                  tau = tau,
                                  m = m, 
                                  graph = graph,
                                  type = type,
                                  type_rational_approximation = type_rational_approximation)
            appr_cov_mat = op$covariance_mesh()
            
            # computing the errors
            L_inf_error = max(abs(true_cov_mat - appr_cov_mat))
            L_2_error = sqrt(as.double(t(graph$mesh$weights)%*%(true_cov_mat - appr_cov_mat)^2%*%graph$mesh$weights))
            
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = L_inf_error
            Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = L_2_error
            }, error = function(err){
              warning(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              print(paste("Error occurred at iteration h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu, "Error:", conditionMessage(err)))
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,1] = NA
              Error[[as.character(h_vector[s])]][[type_vector[i]]][[type_rational_approximation_vector[j]]][[as.character(rho_vector[k])]][[as.character(m_vector[l])]][n,2] = NA
            })
            print(paste("h=", h, ",type=", type, ",type_rational_approximation=", type_rational_approximation, ",rho=", rho, ",m=", m, ",nu=", nu))
          }
        }
      }
    }
  }
}

Error_tadpole_DEF_rho2 = Error
save(Error_tadpole_DEF_rho2, file = here("data_files/Error_tadpole_DEF_rho2.RData"))
```

## Plotting the errors

```{r, class.source = "fold-hide"}
load(here("data_files/Error_tadpole_DEF_rho0.1.RData"))
load(here("data_files/Error_tadpole_DEF_rho0.5.RData"))
load(here("data_files/Error_tadpole_DEF_rho1.RData"))
load(here("data_files/Error_tadpole_DEF_rho2.RData"))

h_vector = c("0.001", "0.005", "0.01", "0.1")
type_rational_approximation_vector = c("chebfun", "brasil", "chebfunLB")
nu_vector = seq(0.1,2.49,by=0.05)[1:29]

n = "0.001"
t = "chebfun"


dat = rbind(
data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,1],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,1],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,1],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,1],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,1],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,1],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,1],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,1],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,1],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,1],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,1],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,1],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,1],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,1],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,1],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,1],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,1]),
           L = rep("Linf", times = 5*29)
),

## L2

data.frame(nu = rep(nu_vector,5), 
           rho = rep(0.1, times = 5*29),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["0"]][1:29,2],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["1"]][1:29,2],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["2"]][1:29,2],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["3"]][1:29,2],
                     Error_tadpole_DEF_rho0.1[[n]][["covariance"]][[t]][["0.1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(0.5, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["0"]][1:29,2],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["1"]][1:29,2],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["2"]][1:29,2],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["3"]][1:29,2],
                     Error_tadpole_DEF_rho0.5[[n]][["covariance"]][[t]][["0.5"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(1, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["0"]][1:29,2],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["1"]][1:29,2],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["2"]][1:29,2],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["3"]][1:29,2],
                     Error_tadpole_DEF_rho1[[n]][["covariance"]][[t]][["1"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
),

data.frame(nu = rep(nu_vector,5), 
           rho = c(rep(2, times = 5*29)),
           m = c(rep(0, 29), rep(1, 29), rep(2, 29), rep(3, 29), rep(4, 29)),
           error = c(Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["0"]][1:29,2],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["1"]][1:29,2],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["2"]][1:29,2],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["3"]][1:29,2],
                     Error_tadpole_DEF_rho2[[n]][["covariance"]][[t]][["2"]][["4"]][1:29,2]),
           L = rep("L2", times = 5*29)
)
)


dat$rho = factor(dat$rho, levels = c("0.1", "0.5", "1", "2"))
dat$L = factor(dat$L, levels = c("L2", "Linf"))
levels(dat$rho) = c("0.1" = latex2exp::TeX("$\\rho = 0.1$"), "0.5" = latex2exp::TeX("$\\rho = 0.5$"), "1" = latex2exp::TeX("$\\rho = 1$"), "2" = latex2exp::TeX("$\\rho = 2$"))
levels(dat$L) = c("L2" = latex2exp::TeX("$L_2(\\Gamma\\times\\Gamma)$ error"), "Linf" = latex2exp::TeX("$L_\\infty(\\Gamma\\times\\Gamma)$ error"))


ggplot(dat, aes(nu, error, colour = as.factor(m))) + 
  facet_grid(L ~ rho, labeller = label_parsed) +
  geom_line() +
  scale_y_log10(n.breaks = 7) +
  scale_x_continuous(n.breaks = 6) +
  theme_bw() +
  theme(panel.spacing = unit(0.3, "cm"), text = element_text(family = "Palatino")) +
  labs(x = bquote(nu ~ "(smoothness parameter)"), y = "error", color = "Order (m)") + ggtitle("Tadpole graph")
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
