---
title: "PeMS 1, covariates and preprocessing"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Go back to the [About page](about.html).


::: {.custom-box}
This vignette processes half of the PeMS dataset, shows some illustrations, and produces [**`pems_repl1_data.RData`**](https://github.com/leninrafaelrierasegura/GWMF/blob/main/data_files/pems_repl1_data.RData), which is a file that contains the graph with data and a covariate defined on the mesh. To see how the data is modeled, go to [pems_repl2.html](pems_repl2.html).
:::


Let us set some global options for all code chunks in this document.


```{r}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
```

Below we load the necessary libraries.

```{r}
library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)

library(ggplot2)
library(cowplot)
library(ggpubr) #annotate_figure()
library(grid) #textGrob()
library(ggmap)

library(viridis)
library(OpenStreetMap)


library(tidyr)
library(sf)

library(here)
library(rmarkdown)
library(grateful) # Cite all loaded packages
```


Below we define the function `captioner()` to generate captions for the figures and the function `gets_summary_parameters()` to extract the summary of the parameters of the model.

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
gets_summary_parameters <- function(fit, model) {
  param_spde <- summary(rspde.result(fit, "field", model, parameterization = "spde"))
  param_matern <- summary(rspde.result(fit, "field", model, parameterization = "matern"))
  param_fixed <- fit$summary.fixed[,1:6]
  marginal.posterior.sigma_e = inla.tmarginal(
    fun = function(x) exp(-x/2), 
    marginal = fit[["internal.marginals.hyperpar"]][["Log precision for the Gaussian observations"]])
  quant.sigma_e <- capture.output({result_tmp <- inla.zmarginal(marginal.posterior.sigma_e)}, file = "/dev/null") 
  quant.sigma_e <- result_tmp
  statistics.sigma_e <- unlist(quant.sigma_e)[c(1,2,3,5,7)]
  mode.sigma_e <- inla.mmarginal(marginal.posterior.sigma_e)
  allparams <- rbind(param_fixed, param_spde, param_matern, c(statistics.sigma_e, mode.sigma_e))
  rownames(allparams)[nrow(allparams)] <- "sigma_e"
  return(allparams)
}
```



<!-- Object `pems_repl` below is available after loading `MetricGraph` library. -->

<!-- ```{r} -->
<!-- # Load the data -->
<!-- pems_data <- pems_repl$data -->
<!-- # Select the locations, which are common for all replicates -->
<!-- PtE_all <- pems_data[1:319, c("edge_number", "distance_on_edge")] -->
<!-- # Select the response variable and transform it to a matrix -->
<!-- Y_all <- matrix(pems_data$y, nrow = 26, ncol = 319, byrow = TRUE) -->
<!-- # Remove locations, those with the same value in the first 13 replicates -->
<!-- all_same <- function(col) {length(unique(col[1:13])) == 1} -->
<!-- cols_to_keep <- apply(Y_all, 2, function(col) !all_same(col)) -->
<!-- Y_loc <- Y_all[, cols_to_keep] -->
<!-- PtE <- PtE_all[cols_to_keep,] -->
<!-- # Check the dimensions -->
<!-- PtE |> dim() -->
<!-- Y_loc |> dim() -->
<!-- ``` -->

<!-- Half of the replicates are used to compute summary statistics (`Y_logstd` and `Y_mean`) and the other half (`Y`) are used to fit the model. -->

<!-- ```{r} -->
<!-- # First half of the replicates for summary statistics -->
<!-- Y_for_summary <- Y_loc[1:13,] -->
<!-- # Reserve the second half for fitting replicate model -->
<!-- Y <- Y_loc[14:26,] -->
<!-- # Compute the log standard deviation and mean -->
<!-- Y_logstd <- apply(Y_for_summary, 2, sd) |> as.vector() |> log() -->
<!-- Y_mean <- apply(Y_for_summary, 2, mean) |> as.vector() -->
<!-- # Check the dimensions -->
<!-- Y |> dim() -->
<!-- ``` -->


********************************************************************************

Folder `data.pems` was downloaded from this [GitHub repository](https://github.com/davidbolin/MetricGraph/tree/7f4c90171c40b9bf7ea29e9b048c93d88477c5d1/examples/data.pems).

```{r}
Lines <- read_sf(here("data_files/data.pems/lines.shp"))
lines <- as_Spatial(Lines)

EtV <- read.csv(here("data_files/data.pems/E.csv"), header = T, row.names = NULL)
PtE <- read.csv(here("data_files/data.pems/PtE.csv"), header = T, row.names = NULL)
PtE[,1] <- PtE[,1] + 1
Y <- read.csv(here("data_files/data.pems/Y.csv"), header = T, row.names = NULL)
Y <- as.matrix(Y[,-1])
edge_length_m <- EtV[,4]
PtE[,2] = PtE[,2]/edge_length_m[PtE[,1]]
```

Matrix `Y` has dimension `r dim(Y)`. Each row of `Y` corresponds to a replicate (`r dim(Y)[1]` replicates in total) and each column corresponds to a location (`r dim(Y)[2]` locations in total) on the network.


```{r}
plot(lines)
PtE |> head(5) |> paged_table()
PtE |> dim()
Y |> as.data.frame() |> head(5) |> paged_table()
Y |> dim()
```


We remove columns in `Y`, those with the same value in the first 13 replicates. We also remove the corresponding rows in `PtE`.

```{r}
all_same <- function(col) {
  length(unique(col[1:13])) == 1
}
cols_to_keep <- apply(Y, 2, function(col) !all_same(col))

Y <- Y[, cols_to_keep]
PtE <- PtE[cols_to_keep,]
PtE |> dim()
Y |> dim()
```


Half of the replicates (`Y_for_summary`) are used to compute `Y_logstd` and `Y_mean`, and the other half (`Y` again ) are used to fit a replicate model in [pems_repl2.html](pems_repl2.html).

```{r}
sampled_numbers <- 1:13
not_sampled_numbers <- 14:26
Y_for_summary <- Y[sampled_numbers,]
Y_logstd <- apply(Y_for_summary, 2, sd) |> as.vector() |> log()
Y_mean <- apply(Y_for_summary, 2, mean) |> as.vector()
Y <- Y[not_sampled_numbers,]
Y |> dim()
```


********************************************************************************


Observe that `Y_logstd` corresponds to $\log$ of the standard deviation at each location.

Below we build the graph using the list element `edges` object from `pems_repl`.

```{r}
# Build the graph
graph <- metric_graph$new(edges = pems_repl$edges, longlat = TRUE)
# Add the observations
graph$add_observations(data = data.frame(y = Y_logstd, 
                                         edge_number = PtE[,1], 
                                         distance_on_edge = PtE[,2]), 
                       normalized = TRUE, 
                       clear_obs = TRUE)
# Build the mesh
graph$build_mesh(h = 0.05) 
```

Below we fit $\log(\mathrm{sd}(s_i))\sim N(\beta_0+u(s_i), \sigma_e^2)$.


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
# Build the model
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "spde")
# Prepare the data for fitting
data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        loc_name = "loc")
# Define the component
cmp_stat <- y ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_stat)
# Fit the model
rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
# Summarize the results
summary(rspde_fit_stat)
gets_summary_parameters(rspde_fit_stat, rspde_model_stat)
```

We now compute the kriging predictor $E(u(s_i)|\log(\mathrm{sd}(s_i)))$ (denoted as `cov` below) and standardize it.

```{r}
# Prediction locations
data_prd_list_mesh <- graph$get_mesh_locations(bru = TRUE, loc = "loc")
# Compute the kriging predictor
y_pred <- predict(rspde_model_stat, cmp_stat, rspde_fit_stat, newdata = data_prd_list_mesh, ~Intercept + field)
cov_almost <- y_pred$pred$mean
# Standardize the kriging predictor of the log standard deviation
cov <- (cov_almost - mean(cov_almost))/sd(cov_almost)
```

Below we plot the results. We first generate two maps, one for zoom level 12 and the other for zoom level 13. 


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>


```{r, class.source = "fold-hide"}
# Extract the Euclidean coordinates of the mesh points
xypoints <- graph$mesh$V
# Extract the range of the coordinates 
x_left <- range(xypoints[,1])[1]
x_right <- range(xypoints[,1])[2]
y_bottom <- range(xypoints[,2])[1]
y_top <- range(xypoints[,2])[2]
# Define coordinates for small windows
coordx_lwr1 <- -121.878
coordx_upr1 <- -121.828
coordy_lwr1 <- 37.315
coordy_upr1 <- 37.365

coordx_lwr2<- -122.075
coordx_upr2 <- -122.025
coordy_lwr2 <- 37.365
coordy_upr2 <- 37.415
# Define the colors for the windows
lower_color <- "darkred"   # Dark purple
upper_color <- "darkblue"  # Yellow
# Set the style of the map
style_vector <- c(c(feature = "road", element = "geometry", visibility = "on"),
                  c("&style=", feature = "poi", element = "labels", visibility = "off"),
                  c("&style=", feature = "road", element = "labels", visibility = "off"),
                  c("&style=", feature = "administrative", element = "labels", visibility = "off"), #name of places
                  c("&style=", feature = "transit", element = "labels", visibility = "off"),
                  c("&style=", feature = "landscape", element = "geometry", visibility = "on"),
                  c("&style=", feature = "water", element = "geometry", visibility = "on")) 
# Obtain the maps
p12 <- ggmap(get_stadiamap(bbox = c(left = x_left, bottom = y_bottom, right = x_right, top = y_top),
              zoom = 12,
              maptype = "stamen_toner_lite", 
              scale =2,
              style = style_vector,
              color = "color")) + xlab(NULL) + ylab(NULL)
p13 <- ggmap(get_stadiamap(bbox = c(left = x_left, bottom = y_bottom, right = x_right, top = y_top),
              zoom = 13,
              maptype = "stamen_toner_lite", 
              scale =2,
              style = style_vector,
              color = "color")) + xlab(NULL) + ylab(NULL)
```

```{r, eval = FALSE, echo = FALSE}
save(p12, p13, file = here("data_files/maps_zoom12and13from_stadia.RData"))
```



Below we plot the field on top of the map. We also add the point values of the standard deviation of the log standard deviation to the graph so we can plot it on top of the map and the field.

```{r, fig.width = 11.2, fig.height = 5.43, class.source = "fold-hide", fig.cap = captioner("Standardized logarithm of the standard deviation of the first part of the data. Points represent the point value at each available location, and edges represent the $\\mathrm{std.cov}(s)$ covariate.")}
# Plot the field on top of the map
f12 <- graph$plot_function(X = cov, 
                          vertex_size = 0, 
                          p = p12,
                          edge_width = 0.5) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")
        ) +
  labs(color = "", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top)
# Plot the field on top of the map
f13 <- graph$plot_function(X = cov, 
                          vertex_size = 0, 
                          p = p13,
                          edge_width = 0.5) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")
        ) +
  labs(color = "", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top)

# We add the point values of standard deviation of the log standard deviation to the graph so we can plot it
standardized_Y_logstd <- (Y_logstd - mean(Y_logstd))/sd(Y_logstd)
graph$add_observations(data = data.frame(y = standardized_Y_logstd, 
                                         edge_number = PtE[,1], 
                                         distance_on_edge = PtE[,2]), 
                       normalized = TRUE, 
                       clear_obs = TRUE)
g12 <- graph$plot(data = "y", vertex_size = 0, p = f12, edge_width = 0, data_size = 1) + 
  labs(color = "", x = "", y = "") + 
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color)
g13 <- graph$plot(data = "y", vertex_size = 0, p = f13, edge_width = 0, data_size = 1) + 
  labs(color = "", x = "", y = "") + 
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) 

r1 <- g13 + xlim(coordx_lwr1, coordx_upr1) + 
                    ylim(coordy_lwr1, coordy_upr1) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

r2 <- g13 + xlim(coordx_lwr2, coordx_upr2) + 
                    ylim(coordy_lwr2, coordy_upr2) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

# Arrange p2 and p3 horizontally
left_col <- plot_grid(r2, r1, labels = NULL, ncol = 1, nrow = 2, rel_heights = c(1,1))

# Combine the top row with p1 in a grid
combined_plot <- plot_grid(left_col, g12, labels = NULL, ncol = 2, rel_widths = c(1,2)) 
final_plot <- annotate_figure(combined_plot, left = textGrob("Latitude", rot = 90, vjust = 1, gp = gpar(cex = 0.8)),
                              bottom = textGrob("Longitude", vjust = -0.5, gp = gpar(cex = 0.8)))
ggsave(here("data_files/stand_log_sigma_3.png"), width = 11.2, height = 5.43, plot = final_plot, dpi = 500)
# Print the combined plot
print(final_plot)
```

We repeat the same process  but now for `Y_mean`. We first add the observations to the graph.

```{r}
graph$add_observations(data = data.frame(y = Y_mean, 
                                         edge_number = PtE[,1], 
                                         distance_on_edge = PtE[,2]),
                       normalized = TRUE, 
                       clear_obs = TRUE)
```

We fit the model $\mathrm{mean}(s_i)\sim N(\beta_0+u(s_i), \sigma_e^2)$.


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
# Build the model
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "spde")
# Prepare the data for fitting
data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        loc_name = "loc")
# Define the component
cmp_stat <- y ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_stat)
# Fit the model
rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
# Summarize the results
summary(rspde_fit_stat)
gets_summary_parameters(rspde_fit_stat, rspde_model_stat)
```


Below we compute the predictions for the mesh and the data location.

```{r}
mean_on_mesh_pred <- predict(rspde_model_stat, cmp_stat, rspde_fit_stat, newdata = data_prd_list_mesh, ~Intercept + field)
cov_for_mean_to_plot <- mean_on_mesh_pred$pred$mean

mean_on_loc_pred <- predict(rspde_model_stat, cmp_stat, rspde_fit_stat, newdata = list(loc = PtE), ~Intercept + field)
cov_for_mean <- mean_on_loc_pred$pred$mean
```

Below we check that the predicted mean is close to the observed mean.

```{r, fig.width = 24, fig.height = 4, fig.cap = captioner("Predicted mean and observed mean.")}
plot(Y_mean, type = "l", col = "darkblue")
lines(cov_for_mean, col = "darkred")
```

Below we plot the results as we did before.


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, fig.width = 11.2, fig.height = 5.43, class.source = "fold-hide", fig.cap = captioner("Average speeds of the first part of the data. Points represent the point value at each available location, and edges represent the $\\mathrm{mean.cov}(s)$ covariate.")}
# Plot the field on top of the map
f12 <- graph$plot_function(X = cov_for_mean_to_plot, 
                          vertex_size = 0, 
                          p = p12,
                          edge_width = 0.5) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")
        ) +
  labs(color = "", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top)
# Plot the field on top of the map
f13 <- graph$plot_function(X = cov_for_mean_to_plot, 
                          vertex_size = 0, 
                          p = p13,
                          edge_width = 0.5) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")
        ) +
  labs(color = "", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top)

g12 <- graph$plot(data = "y", vertex_size = 0, p = f12, edge_width = 0, data_size = 1) + 
  labs(color = "", x = "", y = "") + 
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color)
g13 <- graph$plot(data = "y", vertex_size = 0, p = f13, edge_width = 0, data_size = 1) + 
  labs(color = "", x = "", y = "") + 
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) 

r1 <- g13 + xlim(coordx_lwr1, coordx_upr1) + 
                    ylim(coordy_lwr1, coordy_upr1) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

r2 <- g13 + xlim(coordx_lwr2, coordx_upr2) + 
                    ylim(coordy_lwr2, coordy_upr2) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

# Arrange p2 and p3 horizontally
left_col <- plot_grid(r2, r1, labels = NULL, ncol = 1, nrow = 2, rel_heights = c(1,1))

# Combine the top row with p1 in a grid
combined_plot <- plot_grid(left_col, g12, labels = NULL, ncol = 2, rel_widths = c(1,2)) 
final_plot <- annotate_figure(combined_plot, left = textGrob("Latitude", rot = 90, vjust = 1, gp = gpar(cex = 0.8)),
                              bottom = textGrob("Longitude", vjust = -0.5, gp = gpar(cex = 0.8)))
ggsave(here("data_files/mean_better_3.png"), width = 11.2, height = 5.43, plot = final_plot, dpi = 500)
# Print the combined plot
print(final_plot)
```

We finally add the remaining half of the replicates to the graph and plot one of them.

```{r}
df_rep <- lapply(1:nrow(Y), function(i){data.frame(y = Y[i,],
                                                   mean_value = cov_for_mean,
                                                   edge_number = PtE[,1],
                                                   distance_on_edge = PtE[,2],
                                                   repl = i)})
df_rep <- do.call(rbind, df_rep)

graph$add_observations(data = df_rep, normalized = TRUE, clear_obs = TRUE, group = "repl")
```

Below we plot replicate 14.

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, fig.width = 11.2, fig.height = 5.43, class.source = "fold-hide", fig.cap = captioner("Speed observations (in mph) on the highway network of the city of San Jose in California, recorded on April 3, 2017. The left panels are zoomed-in areas of the panel to the right.")}
g12 <- graph$plot(data = "y", 
                 p = p12,
                 group = 1, 
                 vertex_size = 0, 
                 data_size = 1) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")) +
  labs(color = "", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top) +
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color)

g13 <- graph$plot(data = "y", 
                 p = p13,
                 group = 1, 
                 vertex_size = 0, 
                 data_size = 1) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"), 
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 8),
        plot.margin = unit(-0.4*c(1,0,1,1), "cm")) +
  labs(color = "speed", x = "", y = "") +
  xlim(x_left, x_right) + 
  ylim(y_bottom, y_top) +
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_lwr1, 
           linewidth = 0.4, color = upper_color) +  # Bottom line
  annotate("segment", x = coordx_lwr1, y = coordy_upr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Top line
  annotate("segment", x = coordx_lwr1, y = coordy_lwr1, xend = coordx_lwr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Left line
  annotate("segment", x = coordx_upr1, y = coordy_lwr1, xend = coordx_upr1, yend = coordy_upr1, 
           linewidth = 0.4, color = upper_color) +  # Right line
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_lwr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_upr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_lwr2, y = coordy_lwr2, xend = coordx_lwr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color) +  
  annotate("segment", x = coordx_upr2, y = coordy_lwr2, xend = coordx_upr2, yend = coordy_upr2, 
           linewidth = 0.4, color = lower_color)

r1 <- g13 + xlim(coordx_lwr1, coordx_upr1) + 
                    ylim(coordy_lwr1, coordy_upr1) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

r2 <- g13 + xlim(coordx_lwr2, coordx_upr2) + 
                    ylim(coordy_lwr2, coordy_upr2) + 
  theme(legend.position = "none", 
        plot.margin = unit(-0.2*c(1,1,1,1), "cm"))

# Arrange p2 and p3 horizontally
left_col <- plot_grid(r2, r1, labels = NULL, ncol = 1, nrow = 2, rel_heights = c(1,1))

# Combine the top row with p1 in a grid
combined_plot <- plot_grid(left_col, g12, labels = NULL, ncol = 2, rel_widths = c(1,2)) 
final_plot <- annotate_figure(combined_plot, left = textGrob("Latitude", rot = 90, vjust = 1, gp = gpar(cex = 0.8)),
                              bottom = textGrob("Longitude", vjust = -0.5, gp = gpar(cex = 0.8)))
ggsave(here("data_files/replicate14_3.png"), width = 11.2, height = 5.43, plot = final_plot, dpi = 500)
# Print the combined plot
print(final_plot)
```


```{r}
save(graph, cov, cov_for_mean_to_plot, data_prd_list_mesh, file = here("data_files/pems_repl1_data.RData"))
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```


