---
title: "PeMS replicates"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography:
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```


Go back to the [About page](about.html).

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  #out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)
library(sf)

library(here)
library(rmarkdown)
library(grateful) # Cite all loaded packages

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```

# Preprocessing

## Load the data

Object `pems_repl` is available after loading `MetricGraph` library.

```{r}
pems_data <- pems_repl$data
```

## Partition the data

Half of the replicates are used to compute the summary statistic (`Y_log_sd` and `Y_mean`) and the other half (`Y`) are used to fit the model.

```{r}
# Sample half of the replicates
sampled_numbers <- sample(1:26, 13, replace = FALSE)
not_sampled_numbers <- setdiff(1:26, sampled_numbers)

# Compute the log of the standard deviation at each location
Y_log_sd <- pems_data |> 
  filter(repl %in% sampled_numbers) |> 
  group_by(distance_on_edge, edge_number) |> 
  summarise(log_sd_y = log(sd(y))) |> 
  ungroup()

# Compute the mean at each location
Y_mean <- pems_data |> 
  filter(repl %in% sampled_numbers) |> 
  group_by(distance_on_edge, edge_number) |> 
  summarise(mean_y = mean(y)) |> 
  ungroup()

# Get the remaining replicates for the model
Y <- pems_data |> 
  filter(repl %in% not_sampled_numbers)
```

Observe that `Y_log_sd` corresponds to $\log$ of the standard deviation at each location.

## Biuld the graph

```{r}
# Build the graph
graph <- metric_graph$new(edges = pems_repl$edges, longlat=TRUE)
```

## Covariate for the non-stationary model

```{r}
# Add the observations (log(sd(.)) at each location)
graph$add_observations(data = Y_log_sd, normalized = TRUE, clear_obs = TRUE)

# Build the mesh
graph$build_mesh(h = 0.05) 

# Plot the log of the standard deviation
graph$plot(data = "log_sd_y", vertex_size = 0) + 
  ggtitle("Log(sd(s_i)) using half of the replicates") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

Below we fit $\log(\mathrm{sd}(s_i))\sim N(\beta_0+u(s_i), \sigma_e^2)$.


```{r}
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "matern")


data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        loc_name = "loc")

cmp_stat <- log_sd_y ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_stat)

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

model_for_log_sd_y <- rspde_fit_stat

summary(rspde_fit_stat)
fit.rspde <- rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)
```

We now compute the kriging predictor $E(u(s_i)|\log(\mathrm{sd}(s_i)))$ (denoted as `cov` below) and standardize it.

```{r}
# Define the locations for prediction (mesh points)
data_prd_list_mesh <- graph$get_mesh_locations(bru = TRUE, loc = "loc")

# Compute the kriging predictor
log_sd_y_pred <- predict(rspde_model_stat, cmp_stat, rspde_fit_stat, newdata = data_prd_list_mesh, ~Intercept + field)

# Get the mean of the kriging predictor and standardize it
cov <- log_sd_y_pred$pred$mean
cov <- (cov - mean(cov))/sd(cov)

# Plot the field
graph$plot_function(X = cov, vertex_size = 0) + 
  ggtitle("Standardized kriging predictor of log of sd") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

We use `cov` to build the auxiliary matrices for the non-stationary models.

```{r}
# Non-stationary models for tau and kappa
B.tau = cbind(0, 1, 0, cov, 0)
B.kappa = cbind(0, 0, 1, 0, cov)
```

## Covariate for the mean

```{r}
# Add the observations (mean(.) at each location)
graph$add_observations(data = Y_mean, normalized = TRUE, clear_obs = TRUE)

# Plot the mean
graph$plot(data = "mean_y", vertex_size = 0) + 
  ggtitle("Mean value using half of the replicates") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```


```{r}
# fit
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "matern")


data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        loc_name = "loc")

cmp_stat <- mean_y ~ -1 +
  Intercept(1) +
  field(loc, model = rspde_model_stat)

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

model_for_mean_y <- rspde_fit_stat

summary(rspde_fit_stat)
fit.rspde <- rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)
```


```{r}
# Define list for prediction
data_prd_list <- list()

# Define the locations for prediction
AAA = as.matrix(data.frame(Y_mean$edge_number, Y_mean$distance_on_edge))
colnames(AAA) = c("", "d.e")

# AAA is the observation locations and data_prd_list_mesh$loc is the mesh locations
data_prd_list[["loc"]] = rbind(AAA, data_prd_list_mesh$loc)

# Compute the kriging predictor
mean_y_pred <- predict(rspde_model_stat, cmp_stat, rspde_fit_stat, newdata = data_prd_list, ~Intercept + field)

# Get the mean of the kriging predictor
mean_y_pred_aux <- mean_y_pred$pred$mean

# Get the covariate for the mean
cov_for_mean_aux <- mean_y_pred_aux[1:nrow(Y_mean)]

# Get the covariate for the mean to plot (on the mesh locations)
cov_for_mean_to_plot <- mean_y_pred_aux[(nrow(Y_mean) + 1):length(mean_y_pred_aux)]

# Plot the field
graph$plot_function(X = cov_for_mean_to_plot, vertex_size = 0) + 
  ggtitle("Kriging predictor of the mean") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

## Check the covariate for the mean

```{r, fig.dim = c(20,5)}
plot(Y_mean$mean_y, type = "l", col = "blue")
lines(cov_for_mean_aux, col = "red")
```

## Add the remaining replicates to the graph

We add the remaining half of the replicates to the graph.

```{r}
new_Y <- Y |> 
  mutate(cov_for_mean = rep(cov_for_mean_aux, times = length(not_sampled_numbers)))


graph$add_observations(data = new_Y, normalized = TRUE, clear_obs = TRUE, group = "repl")

data <- graph$get_data()

graph$plot(data = "y", group = 1, vertex_size = 0) + 
  ggtitle("PeMS data, replicate 1") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

## Compute groups for cross-validation


```{r}
# Define aux data frame to compute the distance matrix
aux <- data |> filter(repl == unique(data$repl)[1]) |>
  rename(distance_on_edge = .distance_on_edge, edge_number = .edge_number) |> # Rename the variables (because graph$compute_geodist_PtE() requires so)
  as.data.frame() |> # Transform to a data frame (i.e., remove the metric_graph class)
  dplyr::select(edge_number, distance_on_edge)

# Compute the distance matrix
distmatrix <- graph$compute_geodist_PtE(PtE = aux,
                                             normalized = TRUE,
                                             include_vertices = FALSE)
# Define the distance vector
distance = seq(from = 0, to = 10, by = 0.1)

GROUPS <- list()
for (j in 1:length(distance)) {
  GROUPS[[j]] = list()
  for (i in 1:nrow(aux)) {
    GROUPS[[j]][[i]] <- which(as.vector(distmatrix[i, ]) <= distance[j])
  }
}

NEW_GROUPS <- list()
for (j in 1:length(distance)) {
  my_list <- GROUPS[[j]]
  aux_list <- list()
  for (i in 0:(length(sampled_numbers) - 1)) {
  added_vectors <- lapply(my_list, function(vec) vec + i*nrow(aux))
  aux_list <- c(aux_list, added_vectors)
  }
  NEW_GROUPS[[j]] <- aux_list
}

GROUPS <- NEW_GROUPS
```

## Check the groups

```{r}
indexofinterest <- 326 # Any number between 1 and nrow(data)
pointofinterest <- GROUPS[[20]][[indexofinterest]]
pointofinterest2 <- GROUPS[[50]][[indexofinterest]]
p <- graph$plot(vertex_size = 0) +
  geom_point(data = data, aes(x = .coord_x, y = .coord_y), color = "darkviolet") +
  geom_point(data = data[pointofinterest2, ], aes(x = .coord_x, y = .coord_y), color = "green") +
  geom_point(data = data[pointofinterest, ], aes(x = .coord_x, y = .coord_y), color = "blue") +
  geom_point(data = data[indexofinterest, ], aes(x = .coord_x, y = .coord_y), color = "red") +
  ggtitle("Groups") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino")) +
  coord_fixed()
ggplotly(p)
```


# Modeling

## $\nu = 0.5$

### Stationary model

```{r}
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "spde",
                                       nu = 0.5)

data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        repl = ".all",
                                        loc_name = "loc")
cmp_stat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_stat,
        replicate = data_rspde_bru_stat[["repl"]])

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )
summary(rspde_fit_stat)
fit.rspde <- rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)

rspde_fit_statnu0.5 <- rspde_fit_stat
```


### Non-stationary model


```{r}
rspde_model_nonstat <- rspde.metric_graph(graph,
                                          B.tau = B.tau,
                                          B.kappa =  B.kappa,
                                          parameterization = "spde",
                                          nu = 0.5)

data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                           repl = ".all",
                                           loc_name = "loc")
cmp_nonstat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_nonstat,
        replicate = data_rspde_bru_nonstat[["repl"]])

rspde_fit_nonstat <-
  bru(cmp_nonstat,
      data = data_rspde_bru_nonstat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))


rspde_fit_nonstatnu0.5 <- rspde_fit_nonstat
```

## $\nu = 1.5$

### Stationary model

```{r}
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "spde",
                                       nu = 1.5)

data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        repl = ".all",
                                        loc_name = "loc")

cmp_stat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_stat,
        replicate = data_rspde_bru_stat[["repl"]])

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_stat)
fit.rspde <- rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)

rspde_fit_statnu1.5 <- rspde_fit_stat
```

### Non-stationary model

```{r}
rspde_model_nonstat <- rspde.metric_graph(graph,
                                          B.tau = B.tau,
                                          B.kappa =  B.kappa,
                                          parameterization = "spde",
                                          nu = 1.5)

data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                           repl = ".all",
                                           loc_name = "loc")

cmp_nonstat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_nonstat,
        replicate = data_rspde_bru_nonstat[["repl"]])

rspde_fit_nonstat <-
  bru(cmp_nonstat,
      data = data_rspde_bru_nonstat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))


rspde_fit_nonstatnu1.5 <- rspde_fit_nonstat
```

## $\nu$ estimated

### Stationary model

```{r}
rspde_model_stat <- rspde.metric_graph(graph,
                                       parameterization = "spde")


data_rspde_bru_stat <- graph_data_rspde(rspde_model_stat,
                                        repl = ".all",
                                        loc_name = "loc")

cmp_stat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_stat,
        replicate = data_rspde_bru_stat[["repl"]])

rspde_fit_stat <-
  bru(cmp_stat,
      data = data_rspde_bru_stat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_stat)
fit.rspde <- rspde.result(rspde_fit_stat, "field", rspde_model_stat)
summary(fit.rspde)

rspde_fit_statnuest <- rspde_fit_stat
```

### Non-stationary model

```{r}
rspde_model_nonstat <- rspde.metric_graph(graph,
                                          B.tau = B.tau,
                                          B.kappa =  B.kappa,
                                          parameterization = "spde")

data_rspde_bru_nonstat <- graph_data_rspde(rspde_model_nonstat,
                                           repl = ".all",
                                           loc_name = "loc")

cmp_nonstat <- y ~ -1 +
  Intercept(1) +
  cov_for_mean +
  field(loc, model = rspde_model_nonstat,
        replicate = data_rspde_bru_nonstat[["repl"]])

rspde_fit_nonstat <-
  bru(cmp_nonstat,
      data = data_rspde_bru_nonstat[["data"]],
      family = "gaussian",
      options = list(verbose = FALSE)
  )

summary(rspde_fit_nonstat)
summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))

# rspde_fit_nonstat <-bru_rerun(rspde_fit_nonstat)
# summary(rspde_fit_nonstat)
# summary(rspde.result(rspde_fit_nonstat, "field", rspde_model_nonstat))

rspde_fit_nonstatnuest <- rspde_fit_nonstat
```



# Cross-validation


```{r}
mse.statnu0.5 <- mse.nonstatnu0.5 <- ls.statnu0.5 <- ls.nonstatnu0.5 <- rep(0,length(distance))
mse.statnu1.5 <- mse.nonstatnu1.5 <- ls.statnu1.5 <- ls.nonstatnu1.5 <- rep(0,length(distance))
mse.statnuest <- mse.nonstatnuest <- ls.statnuest <- ls.nonstatnuest <- rep(0,length(distance))

# cross-validation for-loop
for (j in 1:length(distance)) {
  print(j)
  # cross-validation of the stationary model
  cv.statnu0.5 <- inla.group.cv(rspde_fit_statnu0.5, groups = GROUPS[[j]])
  cv.statnu1.5 <- inla.group.cv(rspde_fit_statnu1.5, groups = GROUPS[[j]])
  cv.statnuest <- inla.group.cv(rspde_fit_statnuest, groups = GROUPS[[j]])
  # cross-validation of the nonstationary model
  cv.nonstatnu0.5 <- inla.group.cv(rspde_fit_nonstatnu0.5, groups = GROUPS[[j]])
  cv.nonstatnu1.5 <- inla.group.cv(rspde_fit_nonstatnu1.5, groups = GROUPS[[j]])
  cv.nonstatnuest <- inla.group.cv(rspde_fit_nonstatnuest, groups = GROUPS[[j]])
  # obtain MSE and LS
  mse.statnu0.5[j] <- mean((cv.statnu0.5$mean - data$y)^2)
  mse.statnu1.5[j] <- mean((cv.statnu1.5$mean - data$y)^2)
  mse.statnuest[j] <- mean((cv.statnuest$mean - data$y)^2)
  
  
  mse.nonstatnu0.5[j] <- mean((cv.nonstatnu0.5$mean - data$y)^2)
  mse.nonstatnu1.5[j] <- mean((cv.nonstatnu1.5$mean - data$y)^2)
  mse.nonstatnuest[j] <- mean((cv.nonstatnuest$mean - data$y)^2)
  
  
  ls.statnu0.5[j] <- mean(log(cv.statnu0.5$cv))
  ls.statnu1.5[j] <- mean(log(cv.statnu1.5$cv))
  ls.statnuest[j] <- mean(log(cv.statnuest$cv))
  
  ls.nonstatnu0.5[j] <- mean(log(cv.nonstatnu0.5$cv))
  ls.nonstatnu1.5[j] <- mean(log(cv.nonstatnu1.5$cv))
  ls.nonstatnuest[j] <- mean(log(cv.nonstatnuest$cv))
}

# Create data frames
mse_df <- data.frame(
  distance,
  Statnu0.5 = mse.statnu0.5,
  Nonstatnu0.5 = mse.nonstatnu0.5,
  Statnu1.5 = mse.statnu1.5,
  Nonstatnu1.5 = mse.nonstatnu1.5,
  Statnuest = mse.statnuest,
  Nonstatnuest = mse.nonstatnuest
)

ls_df <- data.frame(
  distance,
  Statnu0.5 = -ls.statnu0.5,
  Nonstatnu0.5 = -ls.nonstatnu0.5,
  Statnu1.5 = -ls.statnu1.5,
  Nonstatnu1.5 = -ls.nonstatnu1.5,
  Statnuest = -ls.statnuest,
  Nonstatnuest = -ls.nonstatnuest
)
```

```{r, fig.dim = c(16,6), fig.cap = "MSE and negative Log-Score as functions of distance (in km) for the stationary (dotdash line, $\\boldsymbol{\\cdot-\\cdot}$) and non-stationary (solid line, $\\boldsymbol{-\\!\\!\\!-\\!\\!\\!-}$)  cases with $\\nu = 0.5$, $\\nu = 1.5$, and $\\nu$ estimated (est)."}

# Convert to long format
mse_long <- mse_df %>%
  pivot_longer(cols = -distance, names_to = "nu", values_to = "MSE")

ls_long <- ls_df %>%
  pivot_longer(cols = -distance, names_to = "nu", values_to = "LogScore")


# Update the label mappings with the new legend title
label_mapping <- c(
  "Statnu0.5" = "0.5", 
  "Nonstatnu0.5" = "0.5", 
  "Statnu1.5" = "1.5", 
  "Nonstatnu1.5" = "1.5", 
  "Statnuest" = "est", 
  "Nonstatnuest" = "est"
)

# Define color and linetype mapping
color_mapping <- c(
  "Statnu0.5" = "blue", 
  "Nonstatnu0.5" = "red", 
  "Statnu1.5" = "black", 
  "Nonstatnu1.5" = "green", 
  "Statnuest" = "darkorange", 
  "Nonstatnuest" = "skyblue"
)

linetype_mapping <- c(
  "Statnu0.5" = "dotdash", 
  "Nonstatnu0.5" = "solid", 
  "Statnu1.5" = "dotdash", 
  "Nonstatnu1.5" = "solid", 
  "Statnuest" = "dotdash", 
  "Nonstatnuest" = "solid"
)

# Plot MSE
mse_plot <- ggplot(mse_long, aes(x = distance, y = MSE, color = nu, linetype = nu)) +
  geom_line(linewidth = 1) +
  labs(y = "MSE", x = "Distance in km") +
  scale_color_manual(values = color_mapping, labels = label_mapping, name = expression(nu)) +
  scale_linetype_manual(values = linetype_mapping, labels = label_mapping, name = expression(nu)) +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))

# Plot negative log-score
ls_plot <- ggplot(ls_long, aes(x = distance, y = LogScore, color = nu, linetype = nu)) +
  geom_line(linewidth = 1) +
  labs(y = "Negative Log-Score", x = "Distance in km") +
  scale_color_manual(values = color_mapping, labels = label_mapping, name = expression(nu)) +
  scale_linetype_manual(values = linetype_mapping, labels = label_mapping, name = expression(nu)) +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))

# Combine plots with a shared legend at the top in a single line
combined_plot <- mse_plot + ls_plot + 
  plot_layout(guides = 'collect') & 
  theme(legend.position = 'right') & 
  guides(color = guide_legend(ncol = 1), linetype = guide_legend(nrow = 1))

# Display combined plot
print(combined_plot)
```

```{r}
save.image(here("data_files/pems_repl_all_data.RData"))
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```


