---
title: "The rol of $\\tau$: Part 2"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Below we set some global options for all code chunks in this document.


```{r}
# Set seed for reproducibility
set.seed(593) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```


Below we load the necessary libraries.

```{r}
library(rSPDE)
library(MetricGraph)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)

library(here)
library(rmarkdown)
# Cite all loaded packages
library(grateful)
```


# Graph construction

Let's build a small graph to illustrate the effect of $\tau$ on the graph.

```{r}
edge1 <- rbind(c(0,0),c(1,0))
edge2 <- rbind(c(0,0),c(0,1))
edge3 <- rbind(c(0,1),c(-1,1))
theta <- seq(from=pi,to=3*pi/2,length.out = 20)
edge4 <- cbind(sin(theta),1+ cos(theta))
edges = list(edge1, edge2, edge3, edge4)

graph <- metric_graph$new(edges = edges)
graph$build_mesh(h = 0.01, continuous = TRUE)
```

```{r, fig.cap = captioner("Edge directions.")}
graph$plot(direction = TRUE) + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino"))
```

Below we generate processes from

\begin{equation}\label{SPDE}\tag{SPDE}
(\kappa^2 - \Delta_{\Gamma})^{\alpha/2}(\tau u) = \Wcal, \quad \text{on $\Gamma$}.
\end{equation}


Below we define the operator and compute the marginal standard deviation $\sigma(s)$ (`est_sigma`) of the process.

```{r}
sigma = 1
range <- 0.5
nu <- 2.5

kappa <- sqrt(8*nu)/range
tau <- sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1
alpha <- nu + 1/2

op <- matern.operators(alpha = alpha, 
                       tau = tau, 
                       kappa = kappa,
                       parameterization = "spde",
                       m = 2, 
                       graph = graph) 


cov <- op$covariance_mesh()
est_sigma <- sqrt(Matrix::diag(cov))
```

Below we construct a non-stationary process $v(s) =\tau(s)u(s)$ (`v`), where $u(s)$ (`u`) is a stationary process and $\tau(s) = \dfrac{1}{\sigma(s)}$ (`tau_func`). Object `u` is a stationary sample from $(\kappa^2 - \Delta_{\Gamma})^{\alpha/2}( u) = \Wcal$ (note that $\tau=1$) and object `v` can be thought of as a non-stationary sample from $(\kappa^2 - \Delta_{\Gamma})^{\alpha/2}(v) = \Wcal$, where $v = \tau u$.

```{r}
normal_sample <- rnorm(dim(cov)[1])
R <- chol(cov)
u = t(R) %*% normal_sample
tau_func <- 1/est_sigma
v = tau_func*u #tau = 1/est_sigma

nonstat_cov <-  Matrix::Diagonal(x = 1/est_sigma)%*%cov%*%Matrix::Diagonal(x = 1/est_sigma)
nonstat_est_sigma <- sqrt(Matrix::diag(nonstat_cov))
```

Observe above that we are getting a sample from a standard normal `normal_sample` and then we multiply it by the Cholesky factor `R` of the covariance matrix `cov` to get a sample from the stationary process `u` with parameters $\alpha =$ `r alpha`, $\kappa =$ `r kappa`, and $\tau =$ `r tau`. We could have also done `simulate(op)` to get a sample from the process `u`. 


Below we are getting a sample from a non-stationary process. Object `nonstat_u` is a non-stationary sample $u$ from $(\kappa^2 - \Delta_{\Gamma})^{\alpha/2}(\tau u) = \Wcal$, where $\tau(s) = \sigma(s)$ and $\kappa(s) = \kappa$. Therefore, $v(s) = \tau(s)u(s)=\sigma(s)u(s)$ and $u(s) = \dfrac{v(s)}{\sigma(s)}$.


The above discussion says that `v` and `nonstat_u`  should coincide. They coincide up to a factor `max(v)/max(nonstat_u)`.

```{r}
B.tau <- cbind(0, 1, 0, log(est_sigma), 0)
B.kappa <- cbind(0, 0, 1, 0, log(kappa)*rep(1, length(normal_sample)))
# Log-regression coefficients
theta <- c(0, 0, 1, 1) 
# Compute the operator
nonstat_op <- rSPDE::spde.matern.operators(graph = graph,
                                      B.tau = B.tau,
                                      B.kappa =  B.kappa,
                                      parameterization = "spde",
                                      theta = theta,
                                      alpha = alpha)$covariance_mesh()
nonstat_R <- chol(nonstat_op)
nonstat_u = t(nonstat_R) %*% normal_sample
nonstat_u_scaled = nonstat_u*max(v)/max(nonstat_u)

nonstat_est_sigma_2 <- sqrt(Matrix::diag(nonstat_op))
nonstat_est_sigma_2_scaled <- sqrt(Matrix::diag(nonstat_op))*max(v)/max(nonstat_u)
```


# Non-stationary fields


:::: {style="display: grid; grid-template-columns: 400px 400px 400px; grid-column-gap: 1px;"}

::: {}


```{r, out.width = "100%", fig.cap = captioner("Non-stationary field constructed as the product of the stationary field and the marginal standard deviation.")}
graph$plot_function(X = v, vertex_size = 0, plotly = TRUE, line_color = "darkgreen", line_width = 3) %>%
  config(mathjax = 'cdn')
```


:::

::: {}


```{r, out.width = "100%", fig.cap = captioner("Non-stationary field constructed by defining non-stationary parameters.")}
graph$plot_function(X = nonstat_u, vertex_size = 0, plotly = TRUE, line_color = "darkgreen", line_width = 3) %>%
  config(mathjax = 'cdn') 
```


:::


::: {}


```{r, out.width = "100%", fig.cap = captioner("Properly scaled non-stationary field constructed by defining non-stationary parameters.")}
graph$plot_function(X = nonstat_u_scaled, vertex_size = 0, plotly = TRUE, line_color = "darkgreen", line_width = 3) %>%
  config(mathjax = 'cdn') 
```


:::

::::


# Marginal standard deviation


:::: {style="display: grid; grid-template-columns: 400px 400px 400px 400px; grid-column-gap: 1px;"}

::: {}


```{r, out.width = "100%", fig.cap = captioner("Marginal standard deviation of the stationary field $u$.")}
graph$plot_function(X = est_sigma, vertex_size = 0, plotly = TRUE, line_color = "darkgreen", line_width = 3) %>%
  config(mathjax = 'cdn')
```


:::

::: {}


```{r, out.width = "100%", fig.cap = captioner("Marginal standard deviation of the non-stationary field $u/\\sigma$.")}
graph$plot_function(X = nonstat_est_sigma, vertex_size = 0, plotly = TRUE, line_color = "blue", line_width = 3) %>%
  config(mathjax = 'cdn') 
```


:::


::: {}


```{r, out.width = "100%", fig.cap = captioner("Marginal standard deviation of the non-stationary field by defining the non-stationary parameter $\\tau = \\sigma$.")}
graph$plot_function(X = nonstat_est_sigma_2, vertex_size = 0, plotly = TRUE, line_color = "red", line_width = 3) %>%
  config(mathjax = 'cdn') 
```


:::


::: {}


```{r, out.width = "100%", fig.cap = captioner("Properly scaled marginal standard deviation of the non-stationary field by defining the non-stationary parameter $\\tau = \\sigma$.")}
graph$plot_function(X = nonstat_est_sigma_2_scaled, vertex_size = 0, plotly = TRUE, line_color = "orange", line_width = 3) %>%
  config(mathjax = 'cdn') 
```


:::

::::

```{r, out.width = "100%", fig.cap = captioner("Difference between the non-stationary field (obtained just by multiplying) and the properly scaled non-stationary field.")}
xx <- v - nonstat_u_scaled
sum(xx)
plot(xx)
```

```{r}
tau
max(v)/max(nonstat_u)
max(nonstat_u)/max(v)
```

Moral of the story: the scaler is $\dfrac{1}{\tau}$, where $\tau$ is the stationary parameter in the initial field

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

