---
title: "Removing consecutive zeros from the speed variable"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
bibliography: 
  - ../references.bib
  - grateful-refs.bib
---

```{r, eval = FALSE, echo = FALSE}
################################################################################
################################################################################
################################################################################
################################################################################
######### DO NOT FORGET TO CHANGE THE TITLE EVERY TIME YOU FIT A MODEL #########
################################################################################
################################################################################
################################################################################
################################################################################
```

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
```

Go back to the [Preprocessing page](../preprocessing.html). This [link](../data_files/README.html) might be useful to keep track of the files created during the preprocessing.

Let us set some global options for all code chunks in this document.


```{r}
knitr::opts_chunk$set(
  message = FALSE,    # Disable messages printed by R code chunks
  warning = FALSE,    # Disable warnings printed by R code chunks
  echo = TRUE,        # Show R code within code chunks in output
  include = TRUE,     # Include both R code and its results in output
  eval = TRUE,       # Evaluate R code chunks
  cache = FALSE,       # Enable caching of R code chunks for faster rendering
  fig.align = "center",
  out.width = "100%",
  retina = 2,
  error = TRUE,
  collapse = FALSE
)
rm(list = ls())
set.seed(1982)
```

# Import libraries

```{r}
# Install R-INLA package
# install.packages("INLA",repos = c(getOption("repos"),INLA ="https://inla.r-inla-download.org/R/testing"), dep = TRUE)
# Update R-INLA package
# inla.upgrade(testing = TRUE)
# Install inlabru package
# remotes::install_github("inlabru-org/inlabru", ref = "devel")
# Install rSPDE package
# remotes::install_github("davidbolin/rspde", ref = "devel")
# Install MetricGraph package
# remotes::install_github("davidbolin/metricgraph", ref = "devel")

library(INLA)
library(inlabru)
library(rSPDE)
library(MetricGraph)

library(plotly)
library(dplyr)
library(tidyr)
library(sf)

library(here) # here() starts from the home directory
library(rmarkdown)
library(grateful) # Cite all loaded packages

rm(list = ls()) # Clear the workspace
set.seed(1982) # Set seed for reproducibility
```

# Load the data

```{r}
# Load the january dataset
load(here("data_files/january.RData"))
```

# Function to remove consecutive zeros from the speed variable

```{r}
# Function to remove consecutive zeros
remove_consecutive_zeros <- function(vec) {
  # Initialize a result vector
  result <- numeric(length(vec))
  # Index for the result vector
  index <- 1
  # Flag to track if the first zero has been encountered
  first_zero <- FALSE
  # Loop through the original vector
  for (i in 1:length(vec)) {
    # If current value is not zero or previous value is not zero, or it's the first zero, add it to result
    if (vec[i] != 0 || (i > 1 && vec[i - 1] != 0) || i == 1) {
      result[index] <- vec[i]
      index <- index + 1
      # Reset first_zero flag if it's the first zero
      if (vec[i] == 0 && !first_zero) {
        first_zero <- TRUE
      }
    } else {
      # Replace consecutive zeros with NA after the first zero
      result[index] <- NA
      index <- index + 1
    }
  }
  # Trim the result vector to remove unused entries
  result <- result[1:(index - 1)]
  return(result)
}
```

# Remove consecutive zeros from the speed variable

```{r}
# Choose the days for the analysis
days <- c(7,14,21,28) # every Thursday of January 2021

# Filter by days and hour of interest
aux <- january %>%
  filter(day %in% days, hour %in% c(13)) %>% # Keep observations between 13:00 and 14:00
  dplyr::select(-PDT, -hour) # Remove PDT and hour variables

# Get the unique buses ID
buses_ID <- unique(aux$ID)

# Remove more than one consecutive zeros from the speed variable and store the data in a new data frame df
df <- aux %>% 
  filter(ID == buses_ID[1], day == days[1]) %>% # Get the first bus in the first day
  arrange(datetime) %>% # Arrange by datetime so that we can remove consecutive zeros
  mutate(speed = remove_consecutive_zeros(speed)) %>% # Remove consecutive zeros
  drop_na(speed) # Drop NAs, if any

for (i in 1:length(buses_ID)) { # Loop through the buses
  for (j in 1:length(days)) { # Loop through the days
    if (i == 1 && j == 1) { # Skip the first bus in the first day, as it has been processed already
      next 
    }
    check <- aux %>% filter(ID == buses_ID[i], day == days[j]) # Get the data for the bus and day
    if(nrow(check) > 0){ # If there is data, process as before and append to df
      tmp <- check %>% 
        arrange(datetime) %>%
        mutate(speed = remove_consecutive_zeros(speed)) %>%
        drop_na(speed)
      df <- rbind(tmp, df)
    }
  }
}

# Change the days from days to 1,2,3, and 4 to make it easier to work with
newdays <- 1:4
df$day <- newdays[match(df$day, days)]

# Save the data corresponding to 7,14,21, and 28 January 2021 during 13:00-14:00 with no consecutive zeros in the speed variable
save(df, file = here("data_files/day7142128hour13noconsecutivezeroes.RData"))
```

# Check the data

```{r}
# Choose a bus ID
IDnumber <- 6697

# Get the data with all zeros for the chosen bus ID 
bus <- january %>% 
  filter(day == 7, ID == IDnumber, hour == 13) %>% 
  arrange(datetime) %>%
  dplyr::select(datetime, speed)

# Plot the speed records for the chosen bus ID
TSstudio::ts_plot(bus,
                  line.mode = "lines+markers",
                  title = paste("Speed records, all zeros, 7 January, 1-2pm, bus ID: ", IDnumber, sep = ""),
                  Xtitle = "Time",
                  Ytitle = "Speed")

# Get the data with no consecutive zeros for the chosen bus ID
bus <- january %>% 
  filter(day == 7, ID == IDnumber, hour == 13) %>% 
  arrange(datetime) %>%
  mutate(speed = remove_consecutive_zeros(speed)) %>%
  drop_na(speed) %>%
  dplyr::select(datetime, speed)

# Plot the speed records for the chosen bus ID
TSstudio::ts_plot(bus,
                  line.mode = "lines+markers",
                  title = paste("Speed records, no consecutive zeros, 7 January, 1-2pm, bus ID: ", IDnumber, sep = ""),
                  Xtitle = "Time",
                  Ytitle = "Speed")

# Get the data with no consecutive zeros for the chosen bus ID (now from the new data frame df to check if it is the same)
bus <- df %>% 
  filter(day == 1, ID == IDnumber) %>%  # Note that day 1 in df corresponds to 7 in january
  arrange(datetime) %>%
  dplyr::select(datetime, speed)

# Plot the speed records for the chosen bus ID
TSstudio::ts_plot(bus,
                  line.mode = "lines+markers",
                  title = paste("Speed records, no consecutive zeros, 7 January, 1-2pm, bus ID: ", IDnumber, sep = ""),
                  Xtitle = "Time",
                  Ytitle = "Speed")
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

