---
title: "Convergence rates for the tadpole graph"
date: "Created: 05-07-2024. Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    # df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```


Go back to the [Convergence Rates page](../convergence_rates.html).

Let us set some global options for all code chunks in this document.


```{r}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

# Import libraries

```{r}
library(rSPDE)
library(MetricGraph)
library(Matrix)

library(dplyr)
library(plotly)
library(scales)
library(patchwork)
library(tidyr)
library(ggplot2)
library(reshape2)
library(sf)

library(here)
library(rmarkdown)
library(knitr)
library(grateful) # Cite all loaded packages

library(latex2exp)
library(plotrix)
```


# Tadpole graph

## Define utility functions

```{r, class.source = "fold-hide"}
# Eigenfunctions for the tadpole graph
tadpole.eig <- function(k,graph){
  x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
  x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 
  
  if(k==0){ 
    f.e1 <- rep(1,length(x1)) 
    f.e2 <- rep(1,length(x2)) 
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    f = list(phi=f1/sqrt(3)) 
    
  } else {
    f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
    f.e2 <- sin(pi*k*x2/2)                  
    
    f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
    
    if((k %% 2)==1){ 
      f = list(phi=f1/sqrt(3)) 
    } else { 
      f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
      f.e2 <- cos(pi*k*x2/2)
      f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
      f <- list(phi=f1,psi=f2/sqrt(3/2))
    }
  }
  
  return(f)
}


# Function to compute the true covariance matrix
gets_true_cov_mat <- function(graph, kappa, tau, alpha, n.overkill){
  Sigma.kl <- matrix(0,nrow = dim(graph$mesh$V)[1],ncol = dim(graph$mesh$V)[1])
  for(i in 0:n.overkill){
    phi <- tadpole.eig(i,graph)$phi
    Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*phi%*%t(phi)
    if(i>0 && (i %% 2)==0){ 
      psi <- tadpole.eig(i,graph)$psi
      Sigma.kl <- Sigma.kl + (1/(kappa^2 + (i*pi/2)^2)^(alpha))*psi%*%t(psi)
    }
    
  }
  Sigma.kl <- Sigma.kl/tau^2
  return(Sigma.kl)
}
```

## Define the graph

```{r, out.width= "50%", fig.cap = captioner("Tadpole graph.")}
edge1 <- rbind(c(0,0), c(1,0))
theta <- seq(from = -pi, to = pi,length.out = 100)
edge2 <- cbind(1+1/pi+cos(theta)/pi, sin(theta)/pi)
edges <- list(edge1, edge2)
graph <- metric_graph$new(edges = edges)
graph$plot() + 
  ggtitle("Tadpole graph") + 
  theme_minimal() + 
  theme(text = element_text(family = "Palatino")) +
  coord_fixed(ratio = 1)
```

## Define the parameters

```{r}
# parameters
h.ok <- 2^-10
type <- "covariance"
type_rational_approximation = "brasil"
rho <- 0.5
#m = 4
sigma <- 1
n.overkill = 1000

# Mesh sizes
h_aux <- seq(5.5, 4.5, by = -1/4)
h_vector <- 2^-h_aux
h_label <- paste0("2^-", h_aux, "")
h_label_latex <- sprintf("$2^{-%f}$", h_aux)

# Alpha values
alpha_aux <- c(6, 7, 8, 9, 12)
alpha_vector <- alpha_aux/8
alpha_label <- paste0(alpha_aux, "/8")
theoretical_rate <- pmin(2*alpha_vector-1/2, 2)
```

## Build the graph with overkill mesh

```{r, out.width= "50%", fig.cap = captioner("Tadpole graph with overkill mesh.")}
graph.ok <- graph$clone()
# Build graph with overkill mesh
graph.ok$build_mesh(h = h.ok)
graph.ok$compute_fem()

graph.ok$plot(mesh = TRUE) +
  ggtitle("Tadpole graph with overkill mesh") +
  theme_minimal() +
  theme(text = element_text(family = "Palatino")) +
  coord_fixed(ratio = 1)
```


```{r}
# Get the overkill mesh locations
loc.ok <- graph.ok$mesh$VtE # or graph.ok$get_mesh_locations()
```


Each `A[[i]]` below is the projection matrix corresponding to the graph with mesh size `h_vector[i]` onto the overkill mesh.

```{r}
# Initialize the list of graphs and the list of projection matrices
graphs <- list()
A <- list()
for(i in 1:length(h_vector)){
  graphs[[i]] <- graph$clone()
  graphs[[i]]$build_mesh(h = h_vector[i])
  A[[i]] <- graphs[[i]]$fem_basis(loc.ok)
}
# Print the dimensions of the projection matrices
print(lapply(A, dim))
```

## Compute the covariance error

```{r, fig.dim = c(16, 20), fig.cap = captioner("Covariance error for the tadpole graph.")}
cov.error.ok.mesh <- matrix(NA, nrow = length(h_vector), ncol = length(alpha_vector))
cov.error.folded <- matrix(NA, nrow = length(h_vector), ncol = length(alpha_vector))

par(family = "Palatino")
layout(matrix(1:(length(alpha_vector)*length(h_vector)),
              nrow = length(alpha_vector),
              byrow = TRUE))

m_values <- c()
for (j in 1:length(alpha_vector)) {
  alpha <- alpha_vector[j]
  fract_alpha <- alpha - floor(alpha)
  nu <- alpha - 0.5
  kappa <- sqrt(8*nu)/rho
  tau <- sqrt(gamma(nu) / (sigma^2 * kappa^(2*nu) * (4*pi)^(1/2) * gamma(nu + 1/2)))  #sigma = 1, d = 1

  Sigma.folded <- gets_true_cov_mat(graph = graph.ok,
                               kappa = kappa,
                               tau = tau,
                               alpha = alpha,
                               n.overkill = n.overkill)

  for (i in 1:length(h_vector)) {
    h <- h_vector[i]
    m <- min(20, 5*ceiling((min(2*alpha - 1/2,2) + 1/2)^2*log(h)^2/(4*pi^2*fract_alpha)))
    m_values <- c(m_values, m)

    Sigma.ok.mesh <- matern.operators(alpha = alpha, # using overkill mesh
                                kappa = kappa,
                                tau = tau,
                                m = m,
                                graph = graph.ok,
                                type = type,
                                type_rational_approximation = type_rational_approximation)$covariance_mesh()


    Sigma <- matern.operators(alpha = alpha,  # this is the approximation
                             kappa = kappa,
                             tau = tau,
                             m = m,
                             graph = graphs[[i]],
                             type = type,
                             type_rational_approximation = type_rational_approximation)$covariance_mesh()

    Sigma.approx <- A[[i]]%*%Sigma%*%t(A[[i]])

    cov.error.ok.mesh[i,j] <- sqrt(as.double(t(graph.ok$mesh$weights)%*%(Sigma.ok.mesh - Sigma.approx)^2%*%graph.ok$mesh$weights))
    cov.error.folded[i,j] <- sqrt(as.double(t(graph.ok$mesh$weights)%*%(Sigma.folded - Sigma.approx)^2%*%graph.ok$mesh$weights))

    aux <- dim(Sigma.approx)[1]
    reordering <- c(1,3:aux,2)
    auxhalf <- reordering[length(reordering) %/% 2 + 1]

    plot(Sigma.folded[auxhalf, reordering], col = "black", type = "l", main = bquote(alpha == .(alpha_label[j]) ~ "," ~ h == .(h_label[i])), xlab = "", ylab = "", lwd = 2, lty = 1)
    lines(Sigma.ok.mesh[auxhalf, reordering], col = "red", lty = 2, lwd = 2)
    lines(Sigma.approx[auxhalf, reordering], col = "blue", lty = 3, lwd = 2)
    legend("topleft", lwd = 2, col = c("black", "red", "blue"), legend = c("true", "ok.mesh", "approx"), lty = c(1,2,3))
  }
}
print(m_values)
```


```{r}
slope_tadpole_folded <- numeric(length(alpha_vector))
slope_tadpole_ok_mesh <- numeric(length(alpha_vector))

# folded
for (u in 1:length(alpha_vector)) {
  slope_tadpole_folded[u] <- coef(lm(log(cov.error.folded[,u]) ~ log(h_vector)))[2]
  }
# ok.mesh
for (u in 1:length(alpha_vector)) {
  slope_tadpole_ok_mesh[u] <- coef(lm(log(cov.error.ok.mesh[,u]) ~ log(h_vector)))[2]
  }

transposed_df <- data.frame(t(data.frame(alpha = alpha_vector, theoretical_rate = theoretical_rate, slope1 = slope_tadpole_folded, slope2 = slope_tadpole_ok_mesh)))
rownames(transposed_df) <- c("alpha", "Theoretical rates", "Folded", "Overkill mesh")
colnames(transposed_df) <- NULL
# Display the transposed data frame
transposed_df |> paged_table()
```

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
loglog_line_equation <- function(x1, y1, slope) {
  b <- log10(y1 / (x1 ^ slope))

  function(x) {
    (x ^ slope) * (10 ^ b)
  }
}

guiding_lines <- matrix(NA, nrow = length(h_vector), ncol = length(alpha_vector))
for (j in 1:length(alpha_vector)) {
  guiding_lines_aux <- matrix(NA, nrow = length(h_vector), ncol = length(h_vector))
  for(k in 1:length(h_vector)){
    point_x1 <- h_vector[k]
    point_y1 <- cov.error.ok.mesh[k, j]
    slope <- theoretical_rate[j]
    line <- loglog_line_equation(x1 = point_x1, y1 = point_y1, slope = slope)
    guiding_lines_aux[,k] <- line(h_vector)
  }
  guiding_lines[,j] <- apply(guiding_lines_aux, 1, mean)
}

# Generate default ggplot2 colors
default_colors <- scales::hue_pal()(ncol(guiding_lines))

# Create the plot_lines list with different colors for each line
plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
  geom_line(data = data.frame(x = h_vector, y = guiding_lines[, i]),
            aes(x = x, y = y), color = default_colors[i], linetype = "dashed", show.legend = FALSE)
})

df <- as.data.frame(cbind(h_vector, cov.error.ok.mesh))
colnames(df) <- c("h_vector", alpha_vector)
df_melted <- melt(df, id.vars = "h_vector", variable.name = "column", value.name = "value")

p <- ggplot() +
  geom_line(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  geom_point(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  plot_lines +
  labs(title = "Covariance error using overkill mesh",
    x = "h",
       y = "Covariance Error",
       color = expression(alpha)) +
  scale_x_log10(breaks = h_vector, labels = round(h_vector,3)) +
  scale_y_log10() +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))
```


<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>

```{r, class.source = "fold-hide"}
guiding_lines <- matrix(NA, nrow = length(h_vector), ncol = length(alpha_vector))
for (j in 1:length(alpha_vector)) {
  guiding_lines_aux <- matrix(NA, nrow = length(h_vector), ncol = length(h_vector))
  for(k in 1:length(h_vector)){
    point_x1 <- h_vector[k]
    point_y1 <- cov.error.folded[k, j]
    slope <- theoretical_rate[j]
    line <- loglog_line_equation(x1 = point_x1, y1 = point_y1, slope = slope)
    guiding_lines_aux[,k] <- line(h_vector)
  }
  guiding_lines[,j] <- apply(guiding_lines_aux, 1, mean)
}

# Generate default ggplot2 colors
default_colors <- scales::hue_pal()(ncol(guiding_lines))

# Create the plot_lines list with different colors for each line
plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
  geom_line(data = data.frame(x = h_vector, y = guiding_lines[, i]),
            aes(x = x, y = y), color = default_colors[i], linetype = "dashed", show.legend = FALSE)
})

df <- as.data.frame(cbind(h_vector, cov.error.folded))
colnames(df) <- c("h_vector", alpha_vector)
df_melted <- melt(df, id.vars = "h_vector", variable.name = "column", value.name = "value")

q <- ggplot() +
  geom_line(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  geom_point(data = df_melted, aes(x = h_vector, y = value, color = column)) +
  plot_lines +
  labs(title = "Covariance error using folded",
    x = "h",
       y = "Covariance Error",
       color = expression(alpha)) +
  scale_x_log10(breaks = h_vector, labels = round(h_vector,3)) +
  scale_y_log10() +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"))
```

```{r, fig.dim = c(12, 5), fig.cap = captioner("Covariance error for the interval graph, using overkill mesh and folded.")}
p + q
```



# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
